---
title: Dynamic Population Mapping (Smartpop) - Part 6
author: Damien C. Jacques
date: '2018-10-08'
slug: dynamic-population-mapping-smartpop-part-6
categories:
  - R
summary: This Notebook details the steps of the data analysis performed in the frame of the Smartpop project -- Dynamic Population WP (part 6).
tags:
  - population
  - mobile phone data
  - statistics
header:
  caption: ''
  image: ''
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE,
                      eval = T)
```

```{r, echo = F}
load("/media/ubuntu/DATA/Data/Smartpop/Matrix/remainder_085km2.Rda")
load("/media/ubuntu/DATA/Data/Smartpop/Matrix/trend_085km2.Rda")
load("/media/ubuntu/DATA/Data/Smartpop/Matrix/WeekHolidays_085km2.Rda")
load("/media/ubuntu/DATA/Data/Smartpop/Matrix/WeekWorkingDays_085km2.Rda")
```

# Build final zone shapefile

```{r, eval = F}
source("/home/ubuntu/Dropbox/Research/scriptR/rasterToPolygons.R")
r <- raster("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2.tif")
shp <- gdal_polygonizeR(r, "/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2.shp")
```

```{r, eval = F}
library(sf)
library(dplyr)

shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2.shp", quiet = T)
shp <- shp[!(shp$DN == 0), ]
shp <- shp[order(shp$DN),]  # needed to correctly cbind matrix data
colnames(shp)[1] <- "ID_Grid" 
shp_dissolve <- shp %>% group_by(ID_Grid) %>% summarize() # remove small polygons
st_write(shp_dissolve, "/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp")
```

```{r, echo = F}
library(sf)

shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
+y_0=5400088.438 +ellps=intl +units=m +no_defs"
```

## Population

```{r, eval = F}
library(raster)
library(rasterVis)
library(classInt)
library(data.table)

zone <- raster("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2.tif")
pop <- raster("/media/ubuntu/DATA/Data/Smartpop/Population/pop_10_100.tif")
pop[zone == 0] <- NA
pop <- pop/1000

# Population density at Voronoi scale in km2
popVoro <- zonal(pop, zone, mean)
popVoro <- popVoro[ -1, ]
colnames(popVoro) <- c("zone", "pop")

fwrite(data.frame(popVoro), "/media/ubuntu/DATA/Data/Smartpop/Population/pop_085km2.csv")

area <- table(zone[,])
area <- as.data.frame(area)
area <- area[-1,]
colnames(area) <- c("zone", "area(km2)")
area$`area(km2)` <- (area$`area(km2)`*50*50) / 1e6


fwrite(area, "/media/ubuntu/DATA/Data/Smartpop/Zone/area_085_inKm2.csv")
```

```{r, echo = F}
library(data.table)

popVoro <- fread("/media/ubuntu/DATA/Data/Smartpop/Population/pop_085km2.csv")
area <- fread("/media/ubuntu/DATA/Data/Smartpop/Zone/area_085_inKm2.csv")  
```


# Spatial cross-validation

## Distance matrix

```{r, eval = F}
library(reshape)
library(geosphere)

shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
+y_0=5400088.438 +ellps=intl +units=m +no_defs"

# Compute zone centroids
shpCentroid <- st_centroid(shp)

# Compute the geodesic for each OD pair
shpCentroid_latlon <- st_transform(shpCentroid, "+init=epsg:4326")

db <- expand.grid.df(data.frame(st_coordinates(shpCentroid_latlon)), 
                     data.frame(st_coordinates(shpCentroid_latlon)))
names(db)[3:4] <- c("X_dest", "Y_dest")

# Vector of distances in meters
db$dist <- distGeo(matrix(c(db$X, db$Y), ncol = 2), 
                   matrix(c(db$X_dest, db$Y_dest), ncol = 2))

# Build OD matrix
db[,c("Origin", "Destination")] <- expand.grid(shpCentroid_latlon$ID_Grid, 
               shpCentroid_latlon$ID_Grid)

# Keep only relevant columns
OD_distance <- db[, c(6, 7, 5)]
rm(db)

# Remove Origin = Destination rows
OD_distance <- OD_distance[-which(OD_distance$Origin == OD_distance$Destination), ]

# Save data
fwrite(OD_distance, "/media/ubuntu/DATA/Data/Smartpop/Zone/OD_distance_085km2.csv")
```

```{r, echo = F}
OD_distance <- fread("/media/ubuntu/DATA/Data/Smartpop/Zone/OD_distance_085km2.csv")
```

The model is trained on a sample of 30% of closest Voronoi polygons from a random polygon and validate using the 70% remaining polygons. 

```{r}
library(Metrics)

popTotal <- sum(popVoro$pop * area$`area(km2)`)

# Set the number of iteration
nIteration <- 1000

# randomly select 100 antennas
set.seed(1)
randomIDs <- popVoro$zone[sample(1:nrow(popVoro), nIteration)]

# Initiate output dataframes
OutputSpatialCV <- data.frame(randomID = randomIDs,
                              rsquared = rep(NA, length(randomIDs)),
                              cor = rep(NA, length(randomIDs)),
                              rmse = rep(NA, length(randomIDs)),
                              intercept = rep(NA, length(randomIDs)),
                              slope = rep(NA, length(randomIDs)))

UncertaintySpatialCV <- data.frame(zone = popVoro$zone,
                                   Iteration1 = rep(NA, length(popVoro$zone)))
for (i in 2:nIteration) {
  UncertaintySpatialCV[, paste0("Iteration", i)] <- NA
}

# Start Spatial CV loop
k = 0
for (randomID in randomIDs) {
  k = k + 1
  # print(k)
  
  # Build training sample
  OD_distance_sub <- OD_distance[OD_distance$Origin == randomID,]
  OD_distance_sub <- OD_distance_sub[order(OD_distance_sub$dist), ][1:994, ]  # 995 = 30% of polygons
  trainingIDs <- c(randomID, OD_distance_sub$Destination)
  trainingWhich <- which(popVoro$zone %in% trainingIDs)
  
  ## Plot training
  # test <- as(shp, "Spatial")
  # test$col <- ifelse(test$ID_Grid %in% trainingIDs, "red", "white")
  # plot(test, col = test$col)
  
  # initiate variable for regression
  y <- as.numeric(log(popVoro$pop[trainingWhich])) 
  x <- as.numeric(log(WeekWorkingDays[trainingWhich,22]/area$`area(km2)`[trainingWhich]))
  # remove log(0) values
  index <- which(x == -Inf | y == -Inf)
  if (length(index) != 0) {
    y <- y[-index]
    x <- x[-index]
  }
  
  # standard log-log regression
  lm.model <- lm(y ~ x) 
  
  # compute prediction data and statistics from lm.model
  x <- as.numeric(log(WeekWorkingDays[-trainingWhich,22]/area$`area(km2)`[-trainingWhich]))
  predData <- exp(lm.model$coefficients[1] + lm.model$coefficients[2]*x)
  predData <- predData * popTotal/sum(predData * area$`area(km2)`[-trainingWhich])
  UncertaintySpatialCV[-trainingWhich, k + 1] <- predData

  OutputSpatialCV$rsquared[OutputSpatialCV$randomID == randomID] <- summary(lm.model)$r.squared
  OutputSpatialCV$cor[OutputSpatialCV$randomID == randomID] <- cor(popVoro$pop[-trainingWhich], predData)
  OutputSpatialCV$rmse[OutputSpatialCV$randomID == randomID] <- rmse(popVoro$pop[-trainingWhich], predData)
  OutputSpatialCV$intercept[OutputSpatialCV$randomID == randomID] <- lm.model$coefficients[1]
  OutputSpatialCV$slope[OutputSpatialCV$randomID == randomID] <- lm.model$coefficients[2]
}
```

```{r, fig.width = 6, fig.height = 4}
par(mfrow = c(1,2))
boxplot(OutputSpatialCV$intercept, ylab = "Alpha")
boxplot(OutputSpatialCV$slope, ylab = "Beta")
par(mfrow = c(1,2))
boxplot(OutputSpatialCV$cor, ylab = "Pearson correlation")
boxplot(OutputSpatialCV$rmse, ylab = "RMSE (inhab./km²)")
```



## Mapping uncertainty


```{r}
library(GADMTools)
map <- gadm.loadCountries(c("BEL"), level = 4, basefile = "./")
plotmap(map)
```

## Load shapefile for subsetting and plotting
```{r}
region <- st_read("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/Belgium_Adm/BEL_adm_shp/BEL_adm3.shp", quiet = T)
region <- st_transform(region, st_crs(shp))

commune <- st_read("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/Belgium_Adm/BEL_adm_shp/BEL_adm4.shp", quiet = T)
commune <- st_transform(commune, st_crs(shp))
```

```{r, fig.width = 8, fig.heigth = 8}
library(matrixStats)
library(cartography)

UncertaintySpatialCV_mean <- rowSds(as.matrix(UncertaintySpatialCV[, -1]), na.rm = T)

shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
+y_0=5400088.438 +ellps=intl +units=m +no_defs"

shp <- cbind(shp, UncertaintySpatialCV_mean)

par(mar = c(0,0,1.5,0))
layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))

# quantization breaks of the rate
bks <- getBreaks(v = shp[["UncertaintySpatialCV_mean"]], method = "quantile", nclass = 5)
bks <- seq(0, 150, 150/5)
bks[6] <- max(shp[["UncertaintySpatialCV_mean"]])

# get a color palette
cols <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20','#bd0026')

choroLayer(x = shp, var = "UncertaintySpatialCV_mean", breaks = bks, col = cols,
           border = "khaki", lwd = 0.001, 
           legend.title.txt = "Uncertainty of population\ndensity prediction (inhab./km²)", 
           legend.pos = 'bottomleft', legend.values.rnd = 0)
layoutLayer(title = paste0("Uncertainty of population density prediction based on spatial cross-validation results (inhab./km²)"), 
            sources = "", north = TRUE, 
            scale = 50, tabtitle = TRUE,
            frame = T,  
            author = "*model based on mobile phone data")

for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
    choroLayer(x = st_crop(shp, region[region$NAME_3 == city,]),
               var = "UncertaintySpatialCV_mean", breaks = bks, col = cols,
               lwd = 0.001, 
               legend.pos = 'n')
    plot(as(commune, "Spatial"), border = "black", lwd = 0.5, add = TRUE)
    layoutLayer(title = paste0("Zoom over ", city), 
                sources = "", scale = 5, tabtitle = TRUE,
                frame = T,  
                author = "")
}

# plot(popVoro$pop, UncertaintySpatialCV_mean, xlab = "Population density (inhab./km²)", 
#      ylab = "Uncertainty (inhab./km²)")
```

We can make similar map with relative uncertainty.

```{r, fig.width = 8, fig.heigth = 8, echo = F}
library(matrixStats)
library(cartography)

UncertaintySpatialCV_mean <- rowSds(as.matrix(UncertaintySpatialCV[, -1]), na.rm = T)
UncertaintySpatialCV_mean <- UncertaintySpatialCV_mean * 100 / popVoro$pop
UncertaintySpatialCV_mean[UncertaintySpatialCV_mean == Inf] <- 0

shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
+y_0=5400088.438 +ellps=intl +units=m +no_defs"

shp <- cbind(shp, UncertaintySpatialCV_mean)

par(mar = c(0,0,1.5,0))
layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))

# quantization breaks of the rate
bks <- getBreaks(v = shp[["UncertaintySpatialCV_mean"]], method = "quantile", nclass = 5)
bks <- seq(0, 100, 100/5)
bks[6] <- max(shp[["UncertaintySpatialCV_mean"]])

# get a color palette
cols <- c('#ffffb2','#fecc5c','#fd8d3c','#f03b20','#bd0026')

choroLayer(x = shp, var = "UncertaintySpatialCV_mean", breaks = bks, col = cols,
           border = "khaki", lwd = 0.001, 
           legend.title.txt = "Uncertainty of population\ndensity prediction (%)", 
           legend.pos = 'bottomleft', legend.values.rnd = 0)
layoutLayer(title = paste0("Uncertainty of population density prediction based on spatial cross-validation results (%)"), 
            sources = "", north = TRUE, 
            scale = 50, tabtitle = TRUE,
            frame = T,  
            author = "*model based on mobile phone data")

for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
    choroLayer(x = st_crop(shp, region[region$NAME_3 == city,]),
               var = "UncertaintySpatialCV_mean", breaks = bks, col = cols,
               lwd = 0.001, 
               legend.pos = 'n')
    plot(as(commune, "Spatial"), border = "black", lwd = 0.5, add = TRUE)
    layoutLayer(title = paste0("Zoom over ", city), 
                sources = "", scale = 5, tabtitle = TRUE,
                frame = T,  
                author = "")
}

# plot(popVoro$pop, UncertaintySpatialCV_mean, xlab = "Population density (inhab./km²)", 
#      ylab = "Uncertainty (inhab./km²)")
```

# Mapping

## Make animated map

```{r, eval = F}
library(cartography)
library(cartogram)
library(animation)
library(sp)

# Plot data
shp <- cbind(shp, WeekHolidays)

ani.options(interval = .10,
            ani.width = 800, 
            ani.height = 800)

saveVideo(movie.name = "/media/ubuntu/DATA/Data/Smartpop/Map/SIM_workingDays.mp4", { 
  for (colname in colnames(shp)[2:5]) {
    bks <- getBreaks(v = shp[[colname]], method = "quantile", nclass = 5)
    # figdim <- getFigDim(x = shp, width = 1000, mar = c(0,0,1.2,0), res = 100)
    
    par(mar = c(0,0,1.5,0))
    layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))
    
    # quantization breaks of the rate
    bks <- getBreaks(v = shp[[colname]], method = "quantile", nclass = 5)
    # get a color palette
    cols <- c('#d7191c','#fdae61','#ffffbf','#abdda4','#2b83ba')
    
    choroLayer(x = shp,var = colname, breaks = bks, col = cols,
               border = "khaki", lwd = 0.01, 
               legend.title.txt = "Population\ndensity", 
               legend.pos = 'bottomleft', legend.values.rnd = 0)
    layoutLayer(title = paste0("Population density in Belgium during working days (", colname, "h)"), 
                sources = "", north = TRUE, 
                scale = 50, tabtitle = TRUE,
                frame = FALSE,  
                author = "*model based on mobile phone data")
    
    for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
      choroLayer(x = st_crop(shp, region[region$NAME_3 == city,]),
                 var = colname, breaks = bks, col = cols,
                 lwd = 0.01, 
                 legend.title.txt = "Population\ndensity", 
                 legend.pos = 'n')
      plot(as(commune, "Spatial"), border = "white", lwd = 0.8, add = TRUE)
      layoutLayer(title = paste0("Zoom over ", city), 
                  sources = "", scale = 5, tabtitle = TRUE,
                  frame = T,  
                  author = "")
    }
  }
})
```

## Risk map


```{r}
library(matrixStats)

datesWeek <- outer(c("Mon.", "Tue.", "Wed.", "Thu.", "Fri.", "Sat.", "Sun."), 0:23, FUN = "paste0")
datesWeek <- t(datesWeek)
dim(datesWeek) <- NULL
hours <- rep(0:23, 7)

PopWorkingDays_WE <- WeekWorkingDays
WeekWorkingDays_WE <- WeekWorkingDays
WeekWorkingDays_WE[is.na(WeekWorkingDays_WE)] <- WeekHolidays[is.na(WeekWorkingDays_WE)]
for (i in 1:168) {
  predData <- exp(mean(OutputSpatialCV$intercept) + mean(OutputSpatialCV$slope) * log(WeekWorkingDays_WE[, i]))
  PopWorkingDays_WE[, i] <- predData * popTotal/sum(predData * area$`area(km2)`)
}

PopWorkingDays_WE_DiffRef_Mean <- (PopWorkingDays_WE[, 22] - rowMeans(PopWorkingDays_WE)) * 100 / PopWorkingDays_WE[, 22] 
PopWorkingDays_WE_DiffRef_CV <- rowSds(PopWorkingDays_WE) * 100 / rowMeans(PopWorkingDays_WE)
PopWorkingDays_WE_DiffRef_which.max_day <- datesWeek[apply(PopWorkingDays_WE,1,which.max)]
PopWorkingDays_WE_DiffRef_which.max_hour <- hours[apply(PopWorkingDays_WE,1,which.max)]

colnames(PopWorkingDays_WE) <- hours
res <- as.data.frame( # sapply returns a list here, so we convert it to a data.frame
    sapply(unique(colnames(PopWorkingDays_WE)), # for each unique column name
       function(col) rowMeans(PopWorkingDays_WE[,colnames(PopWorkingDays_WE) == col]) # calculate row means
    )
  )

PopWorkingDays_WE_DiffRef_which.max_hour <- c(0:23)[apply(res,1,which.max)]
PopWorkingDays_WE_DiffRef_diff_morning_evening <- (rowMeans(res[,11:17]) - rowMeans(res[,22:24])) * 100 / ((rowMeans(res[,11:17]) +rowMeans(res[,22:24]))/2)

#################

PopWeekHolidays_WE <- WeekWorkingDays
for (i in 1:168) {
  predData <- exp(mean(OutputSpatialCV$intercept) + mean(OutputSpatialCV$slope) * log(WeekHolidays[, i]))
  PopWeekHolidays_WE[, i] <- predData * popTotal/sum(predData * area$`area(km2)`)
}

PopWeekHolidays_WE_DiffRef_Mean <- (PopWeekHolidays_WE[, 22] - rowMeans(PopWeekHolidays_WE)) * 100 / PopWeekHolidays_WE[, 22] 
PopWeekHolidays_WE_DiffRef_CV <- rowSds(PopWeekHolidays_WE) * 100 / rowMeans(PopWeekHolidays_WE)
PopWeekHolidays_WE_DiffRef_which.max_day <- datesWeek[apply(PopWeekHolidays_WE,1,which.max)]
PopWeekHolidays_WE_DiffRef_which.max_hour <- hours[apply(PopWeekHolidays_WE,1,which.max)]

colnames(PopWeekHolidays_WE) <- hours
res <- as.data.frame( # sapply returns a list here, so we convert it to a data.frame
    sapply(unique(colnames(PopWeekHolidays_WE)), # for each unique column name
       function(col) rowMeans(PopWeekHolidays_WE[,colnames(PopWeekHolidays_WE) == col]) # calculate row means
    )
  )

PopWeekHolidays_WE_DiffRef_which.max_hour <- c(0:23)[apply(res,1,which.max)]
PopWeekHolidays_WE_DiffRef_diff_morning_evening <- (rowMeans(res[,11:17]) - rowMeans(res[,22:24])) * 100 / ((rowMeans(res[,11:17]) +rowMeans(res[,22:24]))/2)

#################
Pop <- list(PopWorkingDays_WE_DiffRef_Mean, PopWeekHolidays_WE_DiffRef_Mean, 
            PopWorkingDays_WE_DiffRef_CV, PopWeekHolidays_WE_DiffRef_CV, 
            PopWorkingDays_WE_DiffRef_which.max_hour, PopWeekHolidays_WE_DiffRef_which.max_hour,
            PopWorkingDays_WE_DiffRef_diff_morning_evening, PopWeekHolidays_WE_DiffRef_diff_morning_evening)
names(Pop) <- c("PopWorkingDays_WE_DiffRef_Mean", "PopWeekHolidays_WE_DiffRef_Mean", 
                "PopWorkingDays_WE_DiffRef_CV", "PopWeekHolidays_WE_DiffRef_CV",
                "PopWorkingDays_WE_DiffRef_which.max_hour", "PopWeekHolidays_WE_DiffRef_which.max_hour",
                "PopWorkingDays_WE_DiffRef_diff_morning_evening", "PopWeekHolidays_WE_DiffRef_diff_morning_evening")
```

```{r, fig.width = 8, fig.heigth = 8, echo = F}
for (i in 1:2) {
  shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
  st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
  +y_0=5400088.438 +ellps=intl +units=m +no_defs"
  
  shp <- cbind(shp, Pop[[i]])
  colnames(shp)[2] <- names(Pop)[i]
  
  par(mar = c(0,0,1.5,0))
  layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))
  
  # quantization breaks of the rate
  bks <- c(min(shp[[names(Pop)[i]]]), -35, -25, -15, -5, 0, 5, 15, 25, 35, max(shp[[names(Pop)[i]]]))

  # get a color palette
  cols <- c('#67001f','#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac','#053061')
  
  choroLayer(x = shp, var = names(Pop)[i], breaks = bks, col = cols,
             border = "khaki", lwd = 0.001, 
             legend.title.txt = "", 
             legend.pos = 'bottomleft', legend.values.rnd = 0)
  layoutLayer(title = paste0("Relative difference between residential population density and average population density during ", ifelse(i == 1, "working days", "holidays"), " (%)"), 
              sources = "", north = TRUE, 
              scale = 50, tabtitle = TRUE,
              frame = T,  
              author = "*model based on mobile phone data")
  
  for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
      choroLayer(x = st_crop(shp, region[region$NAME_3 == city,]),
                 var = names(Pop)[i], breaks = bks, col = cols,
                 lwd = 0.001, 
                 legend.pos = 'n')
      plot(as(commune, "Spatial"), border = "black", lwd = 0.5, add = TRUE)
      layoutLayer(title = paste0("Zoom over ", city), 
                  sources = "", scale = 5, tabtitle = TRUE,
                  frame = T,  
                  author = "")
  }
}
```
```{r, fig.width = 8, fig.heigth = 8, echo = F}
for (i in 3:4) {
  shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
  st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
  +y_0=5400088.438 +ellps=intl +units=m +no_defs"
  
  shp <- cbind(shp, Pop[[i]])
  colnames(shp)[2] <- names(Pop)[i]
  
  par(mar = c(0,0,1.5,0))
  layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))
  
  # quantization breaks of the rate
  bks <- c(0,5,10,15,20,25,30, max(shp[[names(Pop)[i]]]))

  # get a color palette
  cols <- c('#000000','#4c170a','#951b0b','#e60e04','#ff7400','#ffbf00','#ffff00')
  
  choroLayer(x = shp, var = names(Pop)[i], breaks = bks, col = cols,
             border = "khaki", lwd = 0.001, 
             legend.title.txt = "Coefficient of variation (%)", 
             legend.pos = 'bottomleft', legend.values.rnd = 0)
  layoutLayer(title = paste0("Coefficient of variation of population density during ", ifelse(i == 3, "working days", "holidays"), " (%)"), 
              sources = "", north = TRUE, 
              scale = 50, tabtitle = TRUE,
              frame = T,  
              author = "*model based on mobile phone data")
  
  for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
      choroLayer(x = st_crop(shp, region[region$NAME_3 == city,]),
                 var = names(Pop)[i], breaks = bks, col = cols,
                 lwd = 0.001, 
                 legend.pos = 'n')
      plot(as(commune, "Spatial"), border = "black", lwd = 0.5, add = TRUE)
      layoutLayer(title = paste0("Zoom over ", city), 
                  sources = "", scale = 5, tabtitle = TRUE,
                  frame = T,  
                  author = "")
  }
}
```

```{r, fig.width = 8, fig.heigth = 8, echo = F}
for (i in 5:6) {
  shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
  st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
  +y_0=5400088.438 +ellps=intl +units=m +no_defs"
  
  shp <- cbind(shp, Pop[[i]])
  colnames(shp)[2] <- names(Pop)[i]
  
  par(mar = c(0,0,1.5,0))
  layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))
  
  # quantization breaks of the rate
  bks <- 0:24

  # get a color palette
  cols <- c('#0000ff','#5e25f1','#823de2','#9b52d4','#ae67c5','#bf7bb6','#cd8fa6','#d9a395','#e4b883','#edcc6e','#f5e055','#fcf530','#fff700','#ffe700','#ffd700','#ffc700','#ffb700','#ffa600','#ff9400','#ff8200','#ff6e00','#ff5800','#ff3c00','#ff0000')
  
  choroLayer(x = shp, var = names(Pop)[i], breaks = bks, col = cols,
             border = "khaki", lwd = 0.001, 
             legend.title.txt = "Hour", 
             legend.pos = 'bottomleft', legend.values.rnd = 0)
  layoutLayer(title = paste0("Hour with in average the highest populaiton density during ", ifelse(i == 5, "working days", "holidays")), 
              sources = "", north = TRUE, 
              scale = 50, tabtitle = TRUE,
              frame = T,  
              author = "*model based on mobile phone data")
  
  for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
      choroLayer(x = st_crop(shp, region[region$NAME_3 == city,]),
                 var = names(Pop)[i], breaks = bks, col = cols,
                 lwd = 0.001, 
                 legend.pos = 'n')
      plot(as(commune, "Spatial"), border = "black", lwd = 0.5, add = TRUE)
      layoutLayer(title = paste0("Zoom over ", city), 
                  sources = "", scale = 5, tabtitle = TRUE,
                  frame = T,  
                  author = "")
  }
}
```
```{r, fig.width = 8, fig.heigth = 8, echo = F}
for (i in 7:8) {
  shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
  st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
  +y_0=5400088.438 +ellps=intl +units=m +no_defs"
  
  shp <- cbind(shp, Pop[[i]])
  colnames(shp)[2] <- names(Pop)[i]
  
  par(mar = c(0,0,1.5,0))
  layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))
  
  # quantization breaks of the rate
  bks <- c(min(shp[[names(Pop)[i]]]), -35, -25, -15, -5, 0, 5, 15, 25, 35, max(shp[[names(Pop)[i]]]))

  # get a color palette
  cols <- c('#67001f','#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac','#053061')
  
  choroLayer(x = shp, var = names(Pop)[i], breaks = bks, col = cols,
             border = "khaki", lwd = 0.001, 
             legend.title.txt = "", 
             legend.pos = 'bottomleft', legend.values.rnd = 0)
  layoutLayer(title = paste0("Relative difference between residential population density and average population density during ", ifelse(i == 7, "working days", "holidays"), " (%)"), 
              sources = "", north = TRUE, 
              scale = 50, tabtitle = TRUE,
              frame = T,  
              author = "*model based on mobile phone data")
  
  for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
      choroLayer(x = st_crop(shp, region[region$NAME_3 == city,]),
                 var = names(Pop)[i], breaks = bks, col = cols,
                 lwd = 0.001, 
                 legend.pos = 'n')
      plot(as(commune, "Spatial"), border = "black", lwd = 0.5, add = TRUE)
      layoutLayer(title = paste0("Zoom over ", city), 
                  sources = "", scale = 5, tabtitle = TRUE,
                  frame = T,  
                  author = "")
  }
}
```

```{r, eval = F}
library(raster)

raster_ref <- raster("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2.tif")

shp <- st_read("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m_0.85km2_dissolve.shp", quiet = T)
st_crs(shp) <- "+proj=lcc +lat_1=51.16666723333333 +lat_2=49.8333339 +lat_0=90 +lon_0=4.367486666666666 +x_0=150000.013
+y_0=5400088.438 +ellps=intl +units=m +no_defs"

shp <- cbind(shp, Pop[[1]])
colnames(shp)[2] <- names(Pop)[1]
st_write(shp, "/media/ubuntu/DATA/Data/Smartpop/Risk/PopWorkingDays_WE_DiffRef_Mean.shp")

ext<-extent(raster_ref)
ext<-paste(ext[1], ext[3], ext[2], ext[4])

res<-paste(res(raster_ref)[1], res(raster_ref)[2])

path.in <- "/media/ubuntu/DATA/Data/Smartpop/Risk/PopWorkingDays_WE_DiffRef_Mean.shp"
path.out <- "/media/ubuntu/DATA/Data/Smartpop/Risk/PopWorkingDays_WE_DiffRef_Mean.tif"

attribute<-"PWD_WE_"

command<-'gdal_rasterize'
command<-paste(command, "--config GDAL_CACHEMAX 2000") #Speed-up with more cache (avice: max 1/3 of your total RAM)
command<-paste(command, "-a", attribute) #Identifies an attribute field on the features to be used for a burn in value. The value will be burned into all output bands.
command<-paste(command, "-te", ext) #(GDAL >= 1.8.0) set georeferenced extents. The values must be expressed in georeferenced units. If not specified, the extent of the output file will be the extent of the vector layers.
command<-paste(command, "-tr", res) #(GDAL >= 1.8.0) set target resolution. The values must be expressed in georeferenced units. Both must be positive values.
command<-paste(command, "-ot Int32")
command<-paste(command, path.in)
command<-paste(command, path.out)

system(command)

st_write(commune, "/media/ubuntu/DATA/Data/Smartpop/Risk/commune.shp")

path.in <- "/media/ubuntu/DATA/Data/Smartpop/Risk/commune.shp"
path.out <- "/media/ubuntu/DATA/Data/Smartpop/Risk/commune.tif"

attribute<-"ID_4"

command<-'gdal_rasterize'
command<-paste(command, "--config GDAL_CACHEMAX 2000") #Speed-up with more cache (avice: max 1/3 of your total RAM)
command<-paste(command, "-a", attribute) #Identifies an attribute field on the features to be used for a burn in value. The value will be burned into all output bands.
command<-paste(command, "-te", ext) #(GDAL >= 1.8.0) set georeferenced extents. The values must be expressed in georeferenced units. If not specified, the extent of the output file will be the extent of the vector layers.
command<-paste(command, "-tr", res) #(GDAL >= 1.8.0) set target resolution. The values must be expressed in georeferenced units. Both must be positive values.
command<-paste(command, "-ot Int32")
command<-paste(command, path.in)
command<-paste(command, path.out)

system(command)
```

```{r}
library(raster)
myZonal <- function (x, z, stat, digits = 0, na.rm = TRUE, 
                     ...) { 
  # source: https://stat.ethz.ch/pipermail/r-sig-geo/2013-February/017475.html
  library(data.table)
  fun <- match.fun(stat) 
  vals <- getValues(x) 
  zones <- round(getValues(z), digits = digits) 
  rDT <- data.table(vals, z=zones) 
  setkey(rDT, z) 
  rDT[, lapply(.SD, fun, na.rm = TRUE), by=z] 
} 

zone <- raster("/media/ubuntu/DATA/Data/Smartpop/Risk/commune.tif")
r <- raster("/media/ubuntu/DATA/Data/Smartpop/Risk/PopWorkingDays_WE_DiffRef_Mean.tif")
zstat <- myZonal(r, zone, "mean")
zstat <- data.frame(zstat)
colnames(zstat) <- c("ID_4", "PopWorkingDays_WE_DiffRef_Mean")

commune <- st_read("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/Belgium_Adm/BEL_adm_shp/BEL_adm4.shp", quiet = T)
commune <- st_transform(commune, st_crs(shp))
commune$PopWorkingDays_WE_DiffRef_Mean <- zstat$PopWorkingDays_WE_DiffRef_Mean[-1]
```

```{r, fig.width = 8, fig.heigth = 8, echo = F}
par(mar = c(0,0,1.5,0))
layout(matrix(c(1,1,2,1,1,3, 4, 5,6), 3, 3, byrow = TRUE))

# quantization breaks of the rate
bks <- c(min(commune[["PopWorkingDays_WE_DiffRef_Mean"]]), -12, -8, -4, 0, 4, 8, 12, max(commune[["PopWorkingDays_WE_DiffRef_Mean"]]))

# get a color palette
cols <- c('#b2182b','#d6604d','#f4a582','#fddbc7','#d1e5f0','#92c5de','#4393c3','#2166ac')

choroLayer(x = commune, var = "PopWorkingDays_WE_DiffRef_Mean", breaks = bks, col = cols,
           border = "khaki", lwd = 0.001, 
           legend.title.txt = "", 
           legend.pos = 'bottomleft', legend.values.rnd = 0)
layoutLayer(title = paste0("Relative difference between residential population density and average population density over a working week (%)"), 
            sources = "", north = TRUE, 
            scale = 50, tabtitle = TRUE,
            frame = T,  
            author = "*model based on mobile phone data")
plot(as(commune, "Spatial"), border = "black", lwd = 0.3, add = TRUE)

for (city in c("Brussel", "Antwerpen", "Leuven", "Liège" ,"Namur")) {
    choroLayer(x = st_crop(commune, region[region$NAME_3 == city,]),
               var = "PopWorkingDays_WE_DiffRef_Mean", breaks = bks, col = cols,
               lwd = 0.001, 
               legend.pos = 'n')
    plot(as(commune, "Spatial"), border = "black", lwd = 0.5, add = TRUE)
    layoutLayer(title = paste0("Zoom over ", city), 
                sources = "", scale = 5, tabtitle = TRUE,
                frame = T,  
                author = "")
}

as.data.frame(commune[order(commune$PopWorkingDays_WE_DiffRef_Mean), c("NAME_4", "PopWorkingDays_WE_DiffRef_Mean")])[1:10,-3]
as.data.frame(commune[rev(order(commune$PopWorkingDays_WE_DiffRef_Mean)), c("NAME_4", "PopWorkingDays_WE_DiffRef_Mean")])[1:10,-3]

```


```{r, echo = F, eval = F}
library("FactoMineR")
library("factoextra")

res.pca <- PCA((PopWeekHolidays_WE), scale.unit = F, ncp = 10, graph = TRUE)

fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))

```

