---
title: Map Plots in R
author: Damien C. Jacques
date: '2018-08-27'
slug: map-plots-in-r
summary: "Example of map plots in R."
categories:
  - R
tags:
  - map
  - R
  - plot
  - cartography
output:
  blogdown::html_page:
    toc: yes
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```


### Proportional and Choropleth Symbols Layer

* Author: [Timotée Giraud](https://cv.archives-ouvertes.fr/timothee-giraud)
* [Github](https://gist.github.com/rCarto/006e547288877d04951d086f86476733)
* [Blog](https://rgeomatic.hypotheses.org/1205)
* Packages: `cartography`, `sf`

```{r}
# load libraries
library(sf)
library(cartography)

## Import
# get the path to the shapefile embedded in sf package
shapepath <- system.file("shape/nc.shp", package = "sf")
# import the shapefile
nc <- st_read(shapepath, quiet = T)
# change the projection of the map
nc <- st_transform(nc, 3857)
# get a figure ratio adapted to the basemap
figdim <- getFigDim(x = nc, width = 800, mar = c(0,0,1.2,0), res = 100)

# compute the sudden infant deaths per 1000 births
nc$share <- 100000 * nc$SID74 / nc$BIR74
# quantization breaks of the rate
bks <- getBreaks(v = nc$share, method = "quantile", nclass = 5)
# correct the breaks to use the global rate as limit of class 
global_rate <- sum(nc$SID74) * 100000 / sum(nc$BIR74)
bks[4] <- global_rate
# get a color palette
cols <- carto.pal(pal1 = "green.pal", n1 = 3, pal2 = "wine.pal", n2 = 2)
# set figure margins
par(mar = c(0,0,1.2,0))
# display the North Carolina counties
plot(st_geometry(nc), col = "khaki4", border = "khaki3", 
     bg = "lemonchiffon")
# display circles
propSymbolsChoroLayer(x = nc,var = "BIR74", inches = 0.2,
                      var2 = "share", breaks = bks, col = cols, 
                      legend.var.title.txt = "N. of live births", 
                      legend.var.pos = 'bottomleft', 
                      legend.var2.values.rnd = 0,
                      legend.var2.pos = "topleft", 
                      legend.var2.title.txt = "Sudden infant death syndrome rate*")
# add a title and layout
layoutLayer(title = "Sudden Infant Death Syndrome in North Carolina, 1974-1978", 
            sources = "", north = TRUE, scale = 50,
            theme = "sand.pal", frame = FALSE, 
            author = "*per 100,000 live births. Source: North Carolina SIDS data set")
```

### Contiguous cartogram [(Gastner, Newman 2004)](http://www.pnas.org/content/101/20/7499)

* Author: [Timotée Giraud](https://cv.archives-ouvertes.fr/timothee-giraud)
* [Blog](https://rgeomatic.hypotheses.org/1361)
* Packages: `cartography`, `cartogram`


```{r}
library(sf)
library(cartography)
library(cartogram)

# get the path to the shapefile embedded in sf
# package
shapepath <- system.file("shape/nc.shp", 
                         package = "sf")
# import the shapefile
nc <- sf::st_read(shapepath, quiet = T)
# change the projection of the map
nc <- sf::st_transform(nc, 3857)
# data preparation
# compute the sudden infant deaths per 1000 births
nc$share <- 100000 * nc$SID74 / nc$BIR74

nc <- cartogram_cont(nc, "BIR74")

# quantization breaks of the rate
bks <- getBreaks(v = nc$share, method = "quantile", nclass = 5)
# correct the breaks to use the global rate as limit of class 
global_rate <- sum(nc$SID74) * 100000 / sum(nc$BIR74)
bks[4] <- global_rate
# get a color palette
cols <- carto.pal(pal1 = "green.pal", n1 = 3, 
                  pal2 = "wine.pal", n2 = 2)
## Choropleth layer
# set figure margins and background color
par(mar = c(0,0,1.2,0), bg = "lemonchiffon")
# display the sudden infant deaths per 1000 births
choroLayer(x = nc,var = "share", breaks = bks, col = cols,
           border = "khaki", lwd = 0.5, 
           legend.title.txt = "Sudden infant death syndrome rate*", 
           legend.pos = 'topleft', legend.values.rnd = 0)
# add a title and layout
layoutLayer(title = "Sudden Infant Death Syndrome in North Carolina, 1974-1978", 
            sources = "", north = TRUE, scale = 50,tabtitle = TRUE,
            theme = "sand.pal", frame = FALSE,  
            author = "*per 100,000 live births. Source: North Carolina SIDS data set")
```

### Non-Contiguous cartogram 

* Author: [Timotée Giraud](https://cv.archives-ouvertes.fr/timothee-giraud)
* [Blog](https://rgeomatic.hypotheses.org/1361)
* Packages: `cartography`, `cartogram`, `sf`

```{r}
library(sf)
library(cartography)
library(cartogram)

# get the path to the shapefile embedded in sf
# package
shapepath <- system.file("shape/nc.shp", 
                         package = "sf")
# import the shapefile
nc <- sf::st_read(shapepath, quiet = T)
# change the projection of the map
nc <- sf::st_transform(nc, 3857)
# data preparation
# compute the sudden infant deaths per 1000 births
nc$share <- 100000 * nc$SID74 / nc$BIR74

nc <- cartogram_ncont(nc, "BIR74")

# quantization breaks of the rate
bks <- getBreaks(v = nc$share, method = "quantile", nclass = 5)
# correct the breaks to use the global rate as limit of class 
global_rate <- sum(nc$SID74) * 100000 / sum(nc$BIR74)
bks[4] <- global_rate
# get a color palette
cols <- carto.pal(pal1 = "green.pal", n1 = 3, 
                  pal2 = "wine.pal", n2 = 2)
## Choropleth layer
# set figure margins and background color
par(mar = c(0,0,1.2,0), bg = "lemonchiffon")
# display the sudden infant deaths per 1000 births
choroLayer(x = nc,var = "share", breaks = bks, col = cols,
           border = "khaki", lwd = 0.5, 
           legend.title.txt = "Sudden infant death syndrome rate*", 
           legend.pos = 'topleft', legend.values.rnd = 0)
# add a title and layout
layoutLayer(title = "Sudden Infant Death Syndrome in North Carolina, 1974-1978", 
            sources = "", north = TRUE, scale = 50,tabtitle = TRUE,
            theme = "sand.pal", frame = FALSE,  
            author = "*per 100,000 live births. Source: North Carolina SIDS data set")
```

### Dorling cartogram

* Author: [Timotée Giraud](https://cv.archives-ouvertes.fr/timothee-giraud)
* [Blog](https://rgeomatic.hypotheses.org/1361)
* Packages: `cartography`, `cartogram`, `sf`

```{r}
library(sf)
library(cartography)
library(cartogram)

# get the path to the shapefile embedded in sf
# package
shapepath <- system.file("shape/nc.shp", 
                         package = "sf")
# import the shapefile
nc <- sf::st_read(shapepath, quiet = T)
# change the projection of the map
nc <- sf::st_transform(nc, 3857)
# data preparation
# compute the sudden infant deaths per 1000 births
nc$share <- 100000 * nc$SID74 / nc$BIR74

# Circles' movements weights.
# A weight of 0 prevents any movement. A weight of 1 gives the default movement distance.
w <- 1 - (nc$BIR74 / max(nc$BIR74)) 
nc <- cartogram_dorling(nc, "BIR74",  k = 5, m_weight = w)

# quantization breaks of the rate
bks <- getBreaks(v = nc$share, method = "quantile", nclass = 5)
# correct the breaks to use the global rate as limit of class 
global_rate <- sum(nc$SID74) * 100000 / sum(nc$BIR74)
bks[4] <- global_rate
# get a color palette
cols <- carto.pal(pal1 = "green.pal", n1 = 3, 
                  pal2 = "wine.pal", n2 = 2)
## Choropleth layer
# set figure margins and background color
par(mar = c(0,0,1.2,0), bg = "lemonchiffon")
# display the sudden infant deaths per 1000 births
choroLayer(x = nc,var = "share", breaks = bks, col = cols,
           border = "khaki", lwd = 0.5, 
           legend.title.txt = "Sudden infant death syndrome rate*", 
           legend.pos = 'topleft', legend.values.rnd = 0)
# add a title and layout
layoutLayer(title = "Sudden Infant Death Syndrome in North Carolina, 1974-1978", 
            sources = "", north = TRUE, scale = 50,tabtitle = TRUE,
            theme = "sand.pal", frame = FALSE,  
            author = "*per 100,000 live births. Source: North Carolina SIDS data set")
```

# Mix Cartography

* Author: [Timotée Giraud](https://cv.archives-ouvertes.fr/timothee-giraud)
* [Github](https://gist.github.com/rCarto/ef52aa4e96a7b628956fbf531143ae68)
* [Blog](https://rgeomatic.hypotheses.org/1205)
* Packages: `cartography`, `sf`, `sp`

```{r}
library(cartography)
library(sp)
library(sf)

# Load data
data(nuts2006)

# Save image
sizes <- getFigDim(x = nuts0.spdf, width = 700, mar = c(0,0,0,0), res = 100)

# set margins
opar <- par(mar = c(0,0,0,0))

# Plot basemaps
plot(nuts0.spdf, border = NA, col = NA, bg = "aliceblue")
plot(world.spdf, col  = "#E3DEBF80", border=NA, add=TRUE)
plot(nuts0.spdf, border = "white", col = "#E3DEBF", lwd= 1.1, add=T)


# Plot an OSM Layer
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c('AT', 'CH', "SI", "IT"),]
OSMTILES <- getTiles(x = spdf, type = "osm", zoom = 5, crop = TRUE)
tilesLayer(x = OSMTILES, add=TRUE)


# Plot a choropleth layer
spdf <- nuts2.spdf[substr(nuts2.spdf$id,1,2) =="DE",]
nuts2.df$gdppercap <- nuts2.df$gdppps2008 / nuts2.df$pop2008
choroLayer(spdf = spdf, df = nuts2.df, var = "gdppercap", border = "white", 
           lwd = 0.4, col = carto.pal(pal1 = "sand.pal", n1 = ),
           legend.pos = "n", add= TRUE)


# Plot proportional squares
spdf <- nuts1.spdf[substr(nuts1.spdf$id,1,2) %in% c("BE", "NL", "LU"),]
propSymbolsLayer(spdf = spdf, df = nuts1.df, var = "pop2008", 
                 border = "#7C000C",
                 symbols = "square", inches = 0.2, col = "#ff000080", 
                 legend.pos = "n")


# Plot Penciled region
spdf <- nuts1.spdf[substr(nuts1.spdf$id,1,2) %in% c("TR"),]
spdf@data <- nuts1.df[substr(nuts1.df$id,1,2) %in% c("TR"),]
spdf <- getPencilLayer(x = st_as_sf(spdf), buffer = 50000, size = 200)
typoLayer(x = spdf, var = "id", add=T, col = carto.pal(pal1 = "multi.pal", 12), 
          legend.pos ="n")


# Plot a typologie layer
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c("EE","LT",'LV'),]
typoLayer(spdf = spdf, df = nuts3.df, var = "id", add=T, legend.pos = "n")


# Plot proportional circles
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c("FI"),]
propSymbolsLayer(spdf = spdf, df = nuts3.df, var = "pop2008", 
                 inches = 0.15, col = "#301551", legend.pos = "n", 
                 border = "white")

# Create and plot a grid layer (absolute)
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c("FR"),]
spdf@data <- nuts3.df[match(spdf$id, nuts3.df$id),]
mygrid <- getGridLayer(x = spdf, cellsize = 50000*50000, var = "pop2008")
propSymbolsLayer(x = mygrid, legend.pos = "n", border = "white",
                 var = "pop2008", inches = 0.1, col="darkblue", 
                 add=TRUE)


# Create and plot a grid layer (relative)
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c("ES", "PT"),]
spdf@data <- nuts3.df[match(spdf$id, nuts3.df$id),]
mygrid <- getGridLayer(x = spdf, cellsize = 75000 * 75000, var = "pop2008")
mygrid$densitykm <- mygrid$pop2008 * 1000 * 1000 / mygrid$gridarea 
cols <- carto.pal(pal1 = "wine.pal", n1 = 6)
choroLayer(x = mygrid, var = "densitykm", add=TRUE,
           border = "grey80",col=cols,
           legend.pos = "n", method = "q6")


# Plot a dot density layer
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c("SE", "NO"),]
dotDensityLayer(spdf = spdf, df=nuts3.df,var="pop2008", add = TRUE, 
                col = "grey30",
                n = 100000,  pch = 20, cex = 0.5, legend.pos = "n")

# Plot a proportional links layer
twincities.df <- twincities.df[substr(twincities.df$i,1,2) %in% c("IT") & 
                                 substr(twincities.df$j,1,2) %in% c("IT",'AT', "SI", 
                                                                    "GR", "CH"), ]
twincities.sf <- getLinkLayer(x = nuts2.spdf, df = twincities.df[,1:2])
gradLinkLayer(x = twincities.sf, df = twincities.df,var = "fij",
              legend.pos = "n",
              breaks = c(1,2,4,10), lwd = c(0.5,4,10),
              col = "#92000090", add = TRUE)


# Plot a label layer
spdf <- nuts0.spdf[nuts0.spdf$id %in% c("IS", "CY"),]
df <- data.frame(id = spdf$id, names = c("Cyprus", "Iceland"))
labelLayer(spdf = spdf, df = df, txt =  "names", font = 2, halo = T)


# Plot Discontinuities
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c("PL", "CZ", "SK", "HU"),]
bord <- getBorders(x = spdf)
cols <- carto.pal(pal1 = "blue.pal", n1 = 3, pal2 = "green.pal", n2 = 3)
nuts3.df$gdppercap <- nuts3.df$gdppps2008/nuts3.df$pop2008
choroLayer(spdf = spdf, df = nuts3.df, var = "gdppercap", method = "q6",
           col = cols, add=T, border = NA, legend.pos = F)
discLayer(x = bord, df = nuts3.df,
          var = "gdppercap", col = "red", nclass = 5,
          method = "quantile", threshold = 0.25, sizemin = 1,
          sizemax = 3, type = "rel", 
          legend.pos = "n", add = TRUE)


# Plot a double proportional triangles layer
spdf <- nuts1.spdf[substr(nuts1.spdf$id,1,2) %in% c("IE","UK"),]
propTrianglesLayer(spdf = spdf, df = nuts1.df, var1 = "birth_2008", 
                   var2 = "death_2008", legend.pos = "n", k = 0.075)


# Plot a proportional symbols layer + choro
spdf <- nuts3.spdf[substr(nuts3.spdf$id,1,2) %in% c("GR", "BG", "MK"),]
propSymbolsChoroLayer(spdf = spdf, df = nuts3.df, var2 = "gdppercap", 
                      var = "gdppps2008",add=T, inches = 0.15,
                      col = carto.pal(pal1 = "orange.pal", 8), 
                      legend.var.pos = "n", legend.var2.pos = "n")

# Plot a proportional symbol layer + typo
spdf <- nuts2.spdf[substr(nuts2.spdf$id,1,2) %in% c("DK"),]
nuts2.df$bidon <- 5
propSymbolsTypoLayer(spdf = spdf, df = nuts2.df, var = "bidon", var2 = "id", 
                     add=T, inches = 0.05, col = carto.pal(pal1 = "multi.pal", 5),
                     legend.var.pos = "n", legend.var2.pos = "n")


# smooth Layer
spdf <- nuts2.spdf[substr(nuts2.spdf$id,1,2) %in% c("RO"),]
smoothLayer(spdf = spdf, df = nuts2.df, var = "pop2008", legend.pos = "n",
            col = carto.pal(pal1 = "wine.pal", 3, "turquoise.pal", 3), nclass=6, 
            typefct = "exponential", span = 100000, beta = 3, add=T, mask = spdf, 
            lwd =0.5)



# Plot a layout (sources, scale, text, etc.)
layoutLayer(title = "",
            sources = "Eurostat - 2008, OpenStreetMap & contributors - 2018",  
            author = "T. Giraud & N. Lambert © UMS RIATE - 2018 - cartography v2.1.1", 
            scale = NULL, 
            col = NA, 
            coltitle = "black", 
            frame = FALSE, north = FALSE)  

# plot a scale bar
barscale(size = 500, lwd = 1.5, cex = 0.9)

# plot a north arrow
north(south = T)
```

