---
title: Smartpop - Dynamic Population Mapping
author: Damien C. Jacques
date: '2018-03-27'
slug: smartpop-dynamic-population-mapping
categories:
  - R
tags:
  - population
  - mobile phone data
---


This Notebook details the steps of the analysis of the Dynamic Population WP of the Smartpop project.

## Data Exploration

```{r, message = F}
library(data.table)

Mobile_Phone <- fread("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/MobilePhone/OBE_traffic_per_antenna_201405_201504.csv")

Mobile_Phone
```

### Antenna

How many antennas in the database ?

```{r}
unique_antenna <- unique(Mobile_Phone[, c("CELL_BTS_XCOORD_NUM" ,"CELL_BTS_YCOORD_NUM")])

nrow(unique_antenna)
```

Plot these antennas.

```{r, message = FALSE}
library(rgdal)

antenna_shp <- SpatialPointsDataFrame(unique_antenna[ , 1:2], data.frame(ID = 1:nrow(unique_antenna)))

plot(antenna_shp, pch = "*")
```

Two antennas are mislocated, we will discard them. We have now 5259 antennas.

```{r, message = FALSE}
library(raster)
library(rgeos)

crs(antenna_shp) <- CRS("+init=epsg:31370") # Lambert 72 proj !! NOT EQUAL to EPSG:31300 or EPSG:103300. Made some test to find the good one.

belgium <- getData('GADM', country = 'BE', level = 0)
belgium <- spTransform(belgium, crs(antenna_shp))
belgium <- gBuffer(belgium, width = 5000)  # Buffer to get all antennas close to belgium border

antenna_discarded <- antenna_shp[which(!gIntersects(antenna_shp, belgium, byid=TRUE)), ]
antenna_shp <- antenna_shp[which(gIntersects(antenna_shp, belgium, byid=TRUE)), ]

plot(antenna_shp, pch= "*") +
plot(antenna_discarded, add=T, col="red", pch = "*")  # to be sure the buffer was large enough

nrow(antenna_shp@data)
```

Comparison with the cadastre (from [Walloon Region](http://geoapps.wallonie.be/antennes/#BBOX=-45890.77766488868,367190.77766488865,-12341.35551604438,197672.0853441707)) shows that several antennas from cadaster are not in the db and the other way around. 

```{r, message = FALSE}
cadastre_antenna <- readOGR("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/Cadastre_antenna/antenna_orange_cadastre_31370.shp")
crs(cadastre_antenna) <- CRS("+init=epsg:31370")
cadastre_antenna <- spTransform(cadastre_antenna, crs(antenna_shp))

belgium <- getData('GADM', country='BE', level=4)
liege <- belgium[belgium$NAME_4=="Liège",]
liege <- spTransform(liege, crs(antenna_shp))

plot(liege, main = "antennas from cadastre + database") +
plot(cadastre_antenna, pch="*", col="red", add=T) +
plot(antenna_shp, pch="*", add=T)

# plot(liege, main = "antennas from cadastre") +
# plot(cadastre_antenna, pch="*", col="red", add=T)

# plot(liege, main = "antennas from database") +
# plot(antenna_shp, pch="*", add=T)
```

Leaflet plot for dynamic exploration.

```{r, message = F}
library(leaflet)
library(widgetframe)

antenna_latlon <- spTransform(antenna_shp, CRS("+init=epsg:4326"))
antenna_latlon <- data.frame(lat=coordinates(antenna_latlon)[ , 2], lon=coordinates(antenna_latlon)[ , 1])

cadastre_latlon <- spTransform(cadastre_antenna, CRS("+init=epsg:4326"))
cadastre_latlon <- data.frame(lat=coordinates(cadastre_latlon)[ , 2], lon=coordinates(cadastre_latlon)[ , 1])

l <- leaflet() %>%
  setView(lng = 3.410255, lat = 50.62723, zoom = 12) %>%
  # Base groups
  addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "ESRI World Imagery") %>%
  # Overlay groups
  addCircles(data = antenna_latlon, ~lon, ~lat,  weight = 5,
             color="blue", fillOpacity = 0.9, group = "antenna db") %>%
  addCircles(data = cadastre_latlon, ~lon, ~lat,  weight = 5,
             color="red", fillOpacity = 0.9, group = "antenna cadastre") %>%
  # Layers control
  addLayersControl(
    baseGroups = c("Toner", "OSM (default)", "ESRI World Imagery"),
    overlayGroups = c("antenna db", "antenna cadastre"),
    options = layersControlOptions(collapsed = FALSE)
  )
frameWidget(l)
```

After exploration some cluster of very close antenna exist. Let's further investigate this point.

First, we compute the distance matrix between each antenna.

```{r, eval = FALSE}
library(geosphere)
library(reshape)

dt <- expand.grid.df(antenna_latlon[, c("lon", "lat")], antenna_latlon[, c("lon", "lat")])
names(dt)[3:4] <- c("lon_dest", "lat_dest")
dt$dist <- distGeo(matrix(c(dt$lon, dt$lat), ncol = 2), 
                   matrix(c(dt$lon_dest, dt$lat_dest), ncol = 2))
save(dt, file = "/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/antenna_distance", compress = T)

dt.cad <- expand.grid.df(cadastre_latlon[, c("lon", "lat")], cadastre_latlon[, c("lon", "lat")])
names(dt.cad)[3:4] <- c("lon_dest", "lat_dest")
dt.cad$dist <- distGeo(matrix(c(dt.cad$lon, dt.cad$lat), ncol = 2), 
                   matrix(c(dt.cad$lon_dest, dt.cad$lat_dest), ncol = 2))
save(dt.cad, file = "/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/cadastre_distance", compress = T)
```


Then, we explore the histogram of distances between antennas from the cadastre:

```{r}
load("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/cadastre_distance")

range(dt.cad$dist[dt.cad$dist > 0])
hist(dt.cad$dist[dt.cad$dist > 0 & dt.cad$dist < 50], xlab = "Distance between points", main = "")

length(dt.cad$dist[dt.cad$dist > 0 & dt.cad$dist < 50])/length(dt.cad$dist[dt.cad$dist > 0])
```




```{r, eval = FALSE, echo = FALSE}
load("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/antenna_distance")

hist(dt$dist, xlab = "Distance between points", main = "")
hist(dt$dist[dt$dist < 50 & dt$dist < 50], xlab = "Distance between points", main = "",  breaks = -0.5:50.5)


dt <- dt[dt$dist > 0 & dt$dist < 2, ]
dt <- dt[apply(which(all(x[1:2] == x[3:4]) & all(x[3:4] == x[1:2])),]
-which(dt[, 1:2] == dt[, 3:4] & dt[, 3:4] == dt[, 1:2])

unique(dt$dist[dt$dist < 3])
```

Now, let's plot the number of antennas over time.

