---
title: Dynamic Population Mapping (Smartpop) - Part 2
author: Damien C. Jacques
date: '2018-07-02'
slug: dynamic-population-mapping-smartpop-part-2
categories:
  - R
summary: "This Notebook details the steps of the data analysis carried out for the Smartpop project -- Dynamic Population WP (part 2)."
tags:
  - mobile phone data
  - population
output:
  blogdown::html_page:
    toc: yes
---

Because the number of antennas/BTS is not constant over time (some antennas/BTS are not active during all the hours of the day), we need to build a grid of reference that will be used to estimate the population.  

{{% alert warning %}}
Assumption 4: MP users can only connect to active antenna/BTS. It means that the Voronoi polygons map (corresponding to BTS/antenna coverage) changes every hour because some antenna/BTS are not active during certain hours of the day (i.e. some Voronoi polygons become larger when an antenna/BTS becomes innactive). 
{{% /alert %}}


```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE,
                      eval = FALSE)
```

### Data and packages loading

```{r, eval = F}
library(maptools)
library(rgdal)
library(SDraw)
library(data.table)
library(dplyr)
library(foreach)
library(doMC)
library(raster)
library(sp)

Mobile_Phone <- fread("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/CleanData/MobilePhone.csv")

dates <- unique(Mobile_Phone$DATE_HOUR)

IDbyHOUR <- Mobile_Phone %>%
  group_by(DATE_HOUR) %>%
  summarise(IDs = list(ID))

antenna_shp <- readOGR("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/CleanData/antenna_shp.shp")

belgium <- getData('GADM', country = 'BE', level = 4)
belgium <- spTransform(belgium, crs(antenna_shp))
```

### Reference grid

A reference grid (shapefile of voronoi polygons) is build to have a spatially consistent database with values for each hour in each polygon of reference. All antennas/BTS remaining in the database are used to define this grid. 

```{r, eval = F}
# Voronoi Polygons
all.antenna_sub <- antenna_shp
voro.sub <- voronoi.polygons(all.antenna_sub, belgium)
voro.sub@data <- cbind(voro.sub@data, all.antenna_sub@data$ID)
colnames(voro.sub@data)[4] <- "ID"
writeOGR(voro.sub, "/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m.shp", layer = "zone_50m", driver = "ESRI Shapefile")

# Rasterize Shapefile
path.in <- "/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m.shp"
path.out <- "/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m.tif"
ext <- "23052 21125 295247 243827"
res <- c("50 50")
attribute <- 'ID'

command <- 'gdal_rasterize'
command <- paste(command, "--config COMPRESS LZW") 
command <- paste(command, "-a", attribute) 
command <- paste(command, "-te", ext) 
command <- paste(command, "-tr", res) 
command <- paste(command, "-ot Int16")
command <- paste(command, path.in)
command <- paste(command, path.out)

system(command)
```

### Compute voronoi polygons for each hour

We need to interesect the Voronoi maps of each hour with the reference grid and then disaggregate the MP users of each antenna/BTS in the reference polygons. First let's compute the Voronoi map of each hour.

```{r, eval = F}
sequence <- seq(0, length(dates), 16)

for (k in 1:length(sequence)) {
  print(k)
  registerDoMC(8)
  test <- foreach(i = (sequence[k] + 1):sequence[k + 1]) %dopar% {
    DATE <- dates[i]
    print(DATE)
    all.antenna_sub <- antenna_shp[antenna_shp$ID %in% unlist(IDbyHOUR$IDs[IDbyHOUR$DATE_HOUR == DATE]),]
    voro.sub <- voronoi.polygons(all.antenna_sub, belgium)
    voro.sub@data <- cbind(voro.sub@data, all.antenna_sub@data$ID)
    colnames(voro.sub@data)[4] <- "ID"
    voro.sub$area <- round(voro.sub$area)
    return(voro.sub)
  }
  names(test) <- gsub(" ", "_", dates[(sequence[k] + 1):sequence[k + 1]])
  lapply(seq_along(test), 
         function(i) {writeOGR(test[[i]], 
                      paste0("/media/ubuntu/DATA/Data/Smartpop/Voro/", 
                             names(test)[i], ".shp"),
                      layer = names(test[[i]]), 
                      driver = "ESRI Shapefile")})
}
```

### Rasterize polygons

```{r, eval = F}
filenames <- list.files("/media/ubuntu/DATA/Data/Smartpop/Voro/", pattern = ".shp$")
attribute <- 'ID'

ext <- "23052 21125 295247 243827"
res <- c("50 50")

library(foreach)
library(doMC)
registerDoMC(8)

foreach(i = 1:length(filenames)) %dopar% {
  filename <- filenames[i]
  path.in <- paste0('/media/ubuntu/DATA/Data/Smartpop/Voro/', filename)
  path.out <- paste0('/media/ubuntu/DATA/Data/Smartpop/VoroRaster/', tools::file_path_sans_ext(filename),'_50m.tif')
  
  command <- 'gdal_rasterize'
  command <- paste(command, "--config COMPRESS LZW") 
  command <- paste(command, "-a", attribute) 
  command <- paste(command, "-te", ext) 
  command <- paste(command, "-tr", res) 
  command <- paste(command, "-ot Int16")
  command <- paste(command, path.in)
  command <- paste(command, path.out)
  
  system(command)
}
```

### Zonal stats (tabulate area)

The intersection between each 'hourly Voronoi map' and the reference grid is computed.

```{r, eval = F}
library(raster)

myZonal <- function(r, z, digits = 0, na.rm = TRUE) { 
  vals <- getValues(r) 
  zones <- round(getValues(z), digits = digits)
  rDT <- data.table(vals, z = zones)
  plyr::count(rDT) 
}

z <- raster("/media/ubuntu/DATA/Data/Smartpop/Zone/zone_50m.tif")
filenames <- list.files("/media/ubuntu/DATA/Data/Smartpop/VoroRaster/", pattern = "50m.tif$")
sequence <- seq(0, length(filenames), 4)

for (k in 1:length(sequence)) {
  print(k)
  registerDoMC(4)
  r <- stack(paste0("/media/ubuntu/DATA/Data/Smartpop/VoroRaster/", filenames[(sequence[k] + 1):sequence[k + 1]]))
  
  test <- foreach(i = 1:4) %dopar% {
    Zstat <- myZonal(r = r[[i]], z = z)
    output <- dcast(data = Zstat, z ~ vals, value.var = "freq")
    output[, 2:ncol(output)] <- apply(output[, 2:ncol(output)], 2, function(i) i/sum(i, na.rm = T))
    return(output)
  }

  names(test) <- gsub(" ", "_", dates[(sequence[k] + 1):sequence[k + 1]])
  lapply(seq_along(test), 
         function(i) {fwrite(test[[i]], 
                               paste0("/media/ubuntu/DATA/Data/Smartpop/TabArea/TabArea_", 
                                      names(test)[i], ".csv")
                             )})
}
```

### Transform database based on tabulate area

```{r, eval = F}

library(data.table)
library(raster)

dates <- unique(Mobile_Phone$DATE_HOUR)

for(i in 1:length(dates)) {
  date <- dates[i]
  print(date)
  tabulate <- data.frame(fread(paste0("/media/ubuntu/DATA/Data/Smartpop/TabArea/TabArea_", 
                                      gsub(" ", "_", date), 
                                      ".csv")))
  names(tabulate)[2:ncol(tabulate)] <- substring(names(tabulate)[2:ncol(tabulate)], 2)
  tabulate <- tabulate[-which(tabulate[1, ] == 0), -which(names(tabulate) == 0)]
  Mobile_Phone_Sub <- Mobile_Phone[Mobile_Phone$DATE_HOUR == date, ]
  
  tabulate_copy <- tabulate
  tabulate[ , 2:ncol(tabulate)] <- sweep(tabulate[ , 2:ncol(tabulate)], 
                                         MARGIN = 2, 
                                         Mobile_Phone_Sub$NBR_SIM_BE[order(Mobile_Phone_Sub$ID)],`*`)

  output <- cbind.data.frame(tabulate[, 1], 
                             rowSums(tabulate[, 2:ncol(tabulate)], na.rm = T)) 
  colnames(output) <- c("ID_Grid", "NBR_SIM_BE")
  
  tabulate_copy[ , 2:ncol(tabulate_copy)] <- sweep(tabulate_copy[ , 2:ncol(tabulate_copy)], 
                                         MARGIN = 2, 
                                         Mobile_Phone_Sub$NBR_SIM_ROAMERS[order(Mobile_Phone_Sub$ID)],`*`)
  output$NBR_SIM_ROAMERS <- rowSums(tabulate_copy[, 2:ncol(tabulate_copy)], na.rm = T)
  
  output$DATE_HOUR <- date
  
  fwrite(output, paste0("/media/ubuntu/DATA/Data/Smartpop/DB_new/", gsub(" ", "_", date), '.csv'))
}  


```

### Merge files to get final database

```{r, eval = F}
filenames <- list.files("/media/ubuntu/DATA/Data/Smartpop/DB_new/", full.names = T)

dt <- lapply(filenames, function(x){fread(x)})
db <- rbindlist(dt)
rm(dt)
gc()
db <- db[4045:nrow(db), ]

fwrite(db, "/media/ubuntu/DATA/Data/Smartpop/FinalDB/FinalDB.csv")
```

