---
title: Dynamic Population Mapping (Smartpop) - Part 1
author: Damien C. Jacques
date: '2018-03-27'
categories:
  - R
tags:
  - population
  - mobile phone data
  - statistics
summary: "This Notebook details the steps of the data analysis carried out for the Smartpop project -- Dynamic Population WP."
slug: smartpop-dynamic-population-mapping
output:
  blogdown::html_page:
    toc: yes
---

This Notebook details the steps of the data analysis performed in the frame of the Smartpop project -- Dynamic Population WP.

# Data Cleaning

## Data Loading

```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE,
                      eval = T)
```

```{r, eval = F, message = F}
library(data.table)

Mobile_Phone <- fread("/mobile_phone.csv",  showProgress = F)

Mobile_Phone
```

```{r, echo = F, message = F}
library(data.table)

Mobile_Phone <- fread("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/MobilePhone/OBE_traffic_per_antenna_201405_201504.csv", showProgress = F)

Mobile_Phone
```


## Antennas
### How many antennas / base transceiver stations are in the database ?

```{r}
unique_antenna <- unique(Mobile_Phone[, c("CELL_BTS_XCOORD_NUM" ,"CELL_BTS_YCOORD_NUM")])

nrow(unique_antenna)
```

### Where are they located? 

```{r, message = FALSE}
library(rgdal)

# An ID field is added and will serve as reference for the rest of the analysis
antenna_shp <- SpatialPointsDataFrame(unique_antenna[ , 1:2], 
                                      data.frame(ID = 1:nrow(unique_antenna),
                                                 X = unique_antenna$CELL_BTS_XCOORD_NUM,
                                                 Y = unique_antenna$CELL_BTS_YCOORD_NUM))

par(mar = c(0, 0, 0, 0))
plot(antenna_shp, pch = "*")

# We merge the ID field with the full db
colnames(Mobile_Phone)[2:3] <- c("X", "Y")
Mobile_Phone <- plyr::join(Mobile_Phone, antenna_shp@data, by = c("X", "Y"))
Mobile_Phone <- Mobile_Phone[, -(2:3)]
```

Two BTS/antennas are mislocated (located beyond the Belgium border). They represent less than 0.001% of the total active SIMs in the database. The (2415660,	181728, ID:4877) antenna is very poorly active while the (0, 0, ID:4635) antenna is more active than 25% of the antennas. Removing these antennas should have little impact on the analysis. 

```{r, message = FALSE}
library(raster)
library(rgeos)

crs(antenna_shp) <- CRS("+init=epsg:31370") # Lambert 72 proj !! NOT EQUAL to EPSG:31300 or EPSG:103300. Made some test to find the good one.

belgium <- getData('GADM', country = 'BE', level = 0)
belgium <- spTransform(belgium, crs(antenna_shp))
belgium <- gBuffer(belgium, width = 5000)  # Buffer to get all antennas close to belgium border

antenna_discarded <- antenna_shp[which(!gIntersects(antenna_shp, belgium, byid = TRUE)), ]
```

```{r, message = FALSE, eval = FALSE}
Mobile_Phone_discarded <- Mobile_Phone[Mobile_Phone$ID %in% antenna_discarded$ID,]

sum(Mobile_Phone_discarded$NBR_SIM_BE)*100/sum(as.numeric(Mobile_Phone$NBR_SIM_BE))

#[1] 0.0060622

library(dplyr)

SIMsByAntenna <- Mobile_Phone %>%
  group_by(ID) %>%
  summarise(SUM_BE = sum(NBR_SIM_BE))

quantile(SIMsByAntenna$SUM_BE, seq(0, 1 , 0.05))

#      0%       5%      10%      15%      20%      25%      30%      35%      40%      45%      50%      55%      60% 
#       1    36504    93109   169562   275694   399682   522712   642580   757973   877016   999955  1139617  1283522 
#     65%      70%      75%      80%      85%      90%      95%     100% 
# 1466811  1663672  1878538  2139007  2489182  2949323  3669699 10393655 

Mobile_Phone_discarded %>%
  group_by(ID) %>%
  summarise(SUM_BE = sum(NBR_SIM_BE))

## A tibble: 2 x 2
#     ID SUM_BE
#  <int>  <int>
#1  4635   6122
#2  4877 412115
```
The 2 mislocated antennas are discarded. 5259 antennas remain in the database.

{{% alert warning %}}
Assumption 1: removing the 2 mislocated antennas has no significant impact on the relationship between population and SIMs.
{{% /alert %}}


```{r, message = FALSE}
antenna_shp <- antenna_shp[which(gIntersects(antenna_shp, belgium, byid = TRUE)), ]

par(mar = c(0, 0, 0, 0))
plot(antenna_shp, pch = "*") +
plot(antenna_discarded, add = T, col = "red", pch = "*")  # to be sure the buffer was large enough

nrow(antenna_shp@data)

Mobile_Phone <- Mobile_Phone[-which(Mobile_Phone$ID %in% antenna_discarded$ID),]
```

### Are they correctly located?

Comparison with the cadastre (from [Walloon Region](http://geoapps.wallonie.be/antennes/#BBOX=-45890.77766488868,367190.77766488865,-12341.35551604438,197672.0853441707)) shows that several BTS from the cadastre are not found in the dataset and vice versa. 

```{r, message = FALSE, warning = FALSE}
cadastre_antenna <- readOGR("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/Cadastre_antenna/antenna_orange_cadastre_31370.shp")
crs(cadastre_antenna) <- CRS("+init=epsg:31370")
cadastre_antenna <- spTransform(cadastre_antenna, crs(antenna_shp))

belgium <- getData('GADM', country = 'BE', level = 4)
liege <- belgium[belgium$NAME_4 == "Liège",]
liege <- spTransform(liege, crs(antenna_shp))

plot(liege, main = "zoom on Liège/Luik") +
plot(cadastre_antenna, pch = "*", col = "red", add = T) +
plot(antenna_shp, pch = "*", add = T)
legend("bottomright", c("antennas from cadastre", "antennas from db"), 
       col = c("red", "black"), pch = "*")
```

Leaflet plot for dynamic exploration.

```{r, message = F}
library(leaflet)
library(widgetframe)

antenna_latlon <- spTransform(antenna_shp, CRS("+init=epsg:4326"))
antenna_latlon <- data.frame(lat = coordinates(antenna_latlon)[ , 2], 
                             lon = coordinates(antenna_latlon)[ , 1])

cadastre_latlon <- spTransform(cadastre_antenna, CRS("+init=epsg:4326"))
cadastre_latlon <- data.frame(lat = coordinates(cadastre_latlon)[ , 2], 
                              lon = coordinates(cadastre_latlon)[ , 1])

l <- leaflet() %>%
  setView(lng = 5.578661, lat = 50.633766,, zoom = 12) %>%
  # Base groups
  addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "ESRI World Imagery") %>%
  # Overlay groups
  addCircles(data = antenna_latlon, ~lon, ~lat,  weight = 5,
             color = "blue", fillOpacity = 0.9, group = "antenna db") %>%
  addCircles(data = cadastre_latlon, ~lon, ~lat,  weight = 5,
             color = "red", fillOpacity = 0.9, group = "antenna cadastre") %>%
  # Layers control
  addLayersControl(
    baseGroups = c("Toner", "OSM (default)", "ESRI World Imagery"),
    overlayGroups = c("antenna db", "antenna cadastre"),
    options = layersControlOptions(collapsed = FALSE)
  )
frameWidget(l)
```

### Inter-antenna distance

After exploration, it seems that there are some clusters of very close antenna. Let's further investigate this point.

First, we compute the distance matrix between each antenna.

```{r, eval = FALSE}
library(geosphere)
library(reshape)

dt <- expand.grid.df(antenna_latlon[, c("lon", "lat")], 
                     antenna_latlon[, c("lon", "lat")])
names(dt)[3:4] <- c("lon_dest", "lat_dest")
dt$dist <- distGeo(matrix(c(dt$lon, dt$lat), ncol = 2), 
                   matrix(c(dt$lon_dest, dt$lat_dest), ncol = 2))

dt.cad <- expand.grid.df(cadastre_latlon[, c("lon", "lat")], 
                         cadastre_latlon[, c("lon", "lat")])
names(dt.cad)[3:4] <- c("lon_dest", "lat_dest")
dt.cad$dist <- distGeo(matrix(c(dt.cad$lon, dt.cad$lat), ncol = 2), 
                   matrix(c(dt.cad$lon_dest, dt.cad$lat_dest), ncol = 2))
```
```{r, eval = FALSE, echo = FALSE}
save(dt, file = "/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/antenna_distance", compress = T)
save(dt.cad, file = "/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/cadastre_distance", compress = T)
```
```{r, echo = FALSE}
load("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/cadastre_distance")
```
According to the cadastre, the lowest distance between two antennas is 3.6 m. Less than 0.1 % of the antennas are closer than 1 km.

```{r}
range(dt.cad$dist[dt.cad$dist > 0])

hist(dt.cad$dist[dt.cad$dist > 0]/1000, 
     xlab = "Distance between points (km)", 
     main = "", breaks = 100)
hist(dt.cad$dist[dt.cad$dist > 0 & dt.cad$dist < 1000], 
     xlab = "Distance between points (m)", 
     main = "", breaks = 100)

quantile(dt.cad$dist[dt.cad$dist > 0], c(0.001))
```

In the dataset, 1735 location points are closer than 2 m. In that case, each point probably correspond to antennas from the same BTS.

```{r}
load("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/WorkingDataBase/antenna_distance")

hist(dt$dist/1000, 
     xlab = "Distance between points (km)", 
     main = "",  breaks = 100)
hist(dt$dist[dt$dist > 0 & dt$dist < 1000], 
     xlab = "Distance between points (m)", 
     main = "",  breaks = 100)

CloseAntenna <- dt[dt$dist > 0 & dt$dist < 2,]
length(unique(c(paste(CloseAntenna$lon, CloseAntenna$lat), paste(CloseAntenna$lon_dest, CloseAntenna$lat_dest))))

quantile(dt$dist[dt$dist > 1], c(0.001))
```

```{r, include=FALSE}
rm(dt, dt.cad, antenna_discarded, belgium, liege, cadastre_antenna, CloseAntenna, SIMsByAntenna, Mobile_Phone_discarded, unique_antenna)
gc()
```

{{% alert warning %}}
Assumption 2: Very close location points are antennas from the same BTS OR the same BTS but active at different time of the year. These antennas will be merged. Close antennas (> 2 m, < 1 km) will also be merged but in a further step.
{{% /alert %}}

### Buffer analysis

Two meters buffer around antennas are used to merge the IDs of very close location points.

```{r, warning = F, message = F}
library(rgeos)
library(sp)

# create buffer of 2 m around antennas
buffer1m <- gBuffer(antenna_shp, width = 2)

# find points that falls within the buffers
buffer1mOver  <- over(disaggregate(buffer1m), antenna_shp, returnList = T)

# extract only buffer covered by more than 1 antenna
l <- sapply(buffer1mOver, nrow)
buffer1mOver <- buffer1mOver[which(l > 1)]

# build lookup table to merge close antennas
transposed_buffer1mOver <- lapply(buffer1mOver, function(x) t(x[, 1]))
l <- sapply(transposed_buffer1mOver, length)
transposed_buffer1mOver_3 <-  do.call(rbind, transposed_buffer1mOver[which(l > 2)])
transposed_buffer1mOver_3 <- rbind(transposed_buffer1mOver_3[, 1:2], transposed_buffer1mOver_3[, 3:2])
lookUp <- do.call(rbind, transposed_buffer1mOver[-which(l > 2)])
lookUp <- rbind(lookUp, transposed_buffer1mOver_3)
lookUp <- data.frame(lookUp)
colnames(lookUp) <- c("OldID", "NewID")

# Replace IDs in the full db and antenna shapefile
Mobile_Phone$ID[Mobile_Phone$ID %in% lookUp$OldID]  <- lookUp$NewID[na.omit(match(Mobile_Phone$ID, lookUp$OldID))]
antenna_shp  <- antenna_shp[-which(antenna_shp$ID %in% lookUp$OldID), ] 

# Merge alls identical id (summarise)
library(dplyr)
Mobile_Phone <- Mobile_Phone %>%
                group_by(DATE_HOUR, ID) %>%
                dplyr::summarise(NBR_SIM_BE = sum(as.numeric(NBR_SIM_BE)),
                          NBR_SIM_ROAMERS = sum(as.numeric(NBR_SIM_ROAMERS)),
                          NBR_EVENTS_BE = sum(as.numeric(NBR_EVENTS_BE)),
                          NBR_EVENTS_ROAMERS = sum(as.numeric(NBR_EVENTS_ROAMERS)))

```

The remaining number of antennas in the database is 4375.

```{r}
nrow(antenna_shp)
```

```{r, include=FALSE}
rm(buffer1m, buffer1mOver, lookUp, transposed_buffer1mOver, transposed_buffer1mOver3, transposed_buffer1mOver_3, l)
gc()
```
### Number of BTS over time

Now, let's explore the number of BTS over time.

```{r}
x <- Mobile_Phone %>%                   
                 group_by(DATE_HOUR) %>%          
                 summarise(Unique_Elements = n_distinct(ID))  

plot(1:nrow(x), 
     x$Unique_Elements, 
     type = "l", 
     xlab = "time (hour)", ylab = "Unique antenna")

x <- Mobile_Phone %>%                   
                 group_by(substr(DATE_HOUR, 1, 10)) %>%          
                 summarise(Unique_Elements = n_distinct(ID))

plot(lubridate::ymd(x$`substr(DATE_HOUR, 1, 10)`), 
     x$Unique_Elements, 
     type = "l", 
     xlab = "day", ylab = "Unique antenna")

x <- Mobile_Phone %>%                   
                 group_by(substr(DATE_HOUR, 1, 10)) %>%          
                 summarise(sum_SIM = sum(NBR_SIM_BE))

plot(lubridate::ymd(x$`substr(DATE_HOUR, 1, 10)`), 
     x$sum_SIM, 
     type = "l", 
     xlab = "day", ylab = "N SIM BE")

x <- Mobile_Phone %>%                   
                 group_by(substr(DATE_HOUR, 1, 10)) %>%          
                 summarise(sum_SIM = sum(NBR_EVENTS_BE))

plot(lubridate::ymd(x$`substr(DATE_HOUR, 1, 10)`), 
     x$sum_SIM, 
     type = "l", 
     xlab = "day", ylab = "N EVENTS BE")


x <- Mobile_Phone %>%                   
                 group_by(substr(DATE_HOUR, 1, 10)) %>%          
                 summarise(sum_SIM = sum(NBR_SIM_ROAMERS))

plot(lubridate::ymd(x$`substr(DATE_HOUR, 1, 10)`), 
     x$sum_SIM, 
     type = "l", 
     xlab = "day", ylab = "N SIM ROAMERS")

x <- Mobile_Phone %>%                   
                 group_by(substr(DATE_HOUR, 1, 10)) %>%          
                 summarise(sum_SIM = sum(NBR_EVENTS_ROAMERS))

plot(lubridate::ymd(x$`substr(DATE_HOUR, 1, 10)`), 
     x$sum_SIM, 
     type = "l", 
     xlab = "day", ylab = "N EVENTS ROAMERS")

```

The number of BTS increases over time (trend) with a daily pattern (day / night). For the trend, one can identify three contrasted periods: before 100 first days (around the beginning of August), from 100 to 315 (mid of March) and what remains. 

```{r}
# aggregate SIM activity per day
Mobile_Phone_ByDay <- Mobile_Phone %>%                   
                 group_by(substr(DATE_HOUR, 1, 10), ID) %>%          
                 summarise(SIM_BE = sum(NBR_SIM_BE))

# count the number of active day per antenna
frequencyAntenna <- table(Mobile_Phone_ByDay$ID)
# select antenna active < 365 days in the db
AntennaNotActivetheWholeYear <- names(frequencyAntenna)[frequencyAntenna < 365]
length(AntennaNotActivetheWholeYear)

# subset shapefile into 2 sets of antennas based on AntennaNotActivetheWholeYear
antenna_shp_discard <- antenna_shp[which(antenna_shp$ID %in% AntennaNotActivetheWholeYear), ]
antenna_shp_select <- antenna_shp[-which(antenna_shp$ID %in% AntennaNotActivetheWholeYear), ]
```

```{r, echo = F}
rm(Mobile_Phone_ByDay, antenna_latlon, antenna_shp_discard, antenna_shp_select, cadastre_latlon, x, AntennaNotActivetheWholeYear, frequencyAntenna)
gc()
```

A [shiny app](https://smartpop.shinyapps.io/ExploreAntenna/) is developped to check, one by one, each BTS with less than 365 active days (609). 

```{r, message = F, eval = F}
library(leaflet)

# make shapefile readable in leaflet
antenna_discard_latlon <- spTransform(antenna_shp_discard, CRS("+init=epsg:4326"))
antenna_discard_latlon <- data.frame(lat = coordinates(antenna_discard_latlon)[ , 2],
                                     lon = coordinates(antenna_discard_latlon)[ , 1],
                                     ID = as.character(antenna_shp_discard$ID))

antenna_select_latlon <- spTransform(antenna_shp_select, CRS("+init=epsg:4326"))
antenna_select_latlon <- data.frame(lat = coordinates(antenna_select_latlon)[ , 2],
                                    lon = coordinates(antenna_select_latlon)[ , 1],
                                     ID = as.character(antenna_shp_select$ID))

## make the shiny app

library(shiny)
# make a copy of Mobile_Phone_ByDay
Mobile_Phone_ByDay_comp <- Mobile_Phone_ByDay
# Store all date (using an antenna active the 365 days)
dates <- lubridate::ymd(Mobile_Phone_ByDay$`substr(DATE_HOUR, 1, 10)`[Mobile_Phone_ByDay$ID==1211])


server <- function(input, output) {
  # Leaflet map with markers
  output$map <- renderLeaflet({
    leaflet() %>%
      setView(lng = 5.578661, lat = 50.633766, zoom = 12) %>%
      addProviderTiles(providers$Stamen.Toner, group = "Toner") %>%
      addCircleMarkers(data = antenna_select_latlon, ~lon, ~lat, layerId= ~ID, popup = ~ID, radius = 4,
                       color="blue", fillOpacity = 0.9,  stroke = TRUE) %>%
      addCircleMarkers(data = antenna_discard_latlon, ~lon, ~lat, layerId= ~ID, popup = ~ID, radius = 4,
                       color="red", fillOpacity = 0.9,  stroke = TRUE)
  })
  
  # Generate data in reactive for time series 1
  plot_data <- reactive({
    ID <- input$map_marker_click$id
    if(is.null(ID)){ID=800}
    Mobile_Phone_ByDay[Mobile_Phone_ByDay$ID %in% ID,]
  })
  
  # Generate data in reactive for time series 2
  plot_data_comp <- reactive({
    ID_comp <- input$location
    
    if(is.null(ID_comp)){ID_comp=800}
    Mobile_Phone_ByDay_comp[Mobile_Phone_ByDay_comp$ID %in% ID_comp,]
  })
  
  # Zoom on selected antenna using the drop-down menu
  observeEvent(input$location, {
    p2 <- antenna_discard_latlon[antenna_discard_latlon$ID == input$location,]
    leafletProxy("map") %>% 
      setView(lng=p2$lon, lat=p2$lat, zoom=14) %>% 
      addCircleMarkers(p2$lon, p2$lat, radius=8, color="black", 
                       fillColor="red", fillOpacity=1, opacity=1, 
                       stroke=TRUE, layerId="Selected")
    
  })
  
  # Make a plot depending on the selected antennas
  output$plot=renderPlot({
    plot(dates, rep(NA, 365), ylim = c(0, 15000), ylab = "N SIM BE")
    lines(lubridate::ymd(plot_data_comp()$`substr(DATE_HOUR, 1, 10)`),  
          plot_data_comp()$SIM_BE, 
          col = "red")
    lines(lubridate::ymd(plot_data()$`substr(DATE_HOUR, 1, 10)`),
          plot_data()$SIM_BE, 
          col ="black")
    legend("topright", col = c("red", "black"), lty = 1,
           c("active < 365 days", "active <= 365 days"),
           bty = "n")
  })
  
  # Generate text with the number of active days for the antenna selected with the drop-down menu
  output$n_days <- renderText({ 
    paste("N active days:", frequencyAntenna[which(names(frequencyAntenna) == input$location)])
  })
}

# Make user interface
ui <- fluidPage(
  br(),
  fluidRow(
    column(4, h2("Explore Antennas (SMARTPOP)"), h3("Click on antenna for comparison (black line)")),
    column(4, selectInput("location", "Antenna active < 365 days (red line)", 
                          c("", as.character(antenna_discard_latlon$ID)), 
                          selected="", multiple=F, width="100%")),
    column(4, textOutput("n_days"))
  ),
  fluidRow(
    column(4, leafletOutput("map")),
    column(8, plotOutput("plot"))
  ),
  br()
)

# Launch the app
shinyApp(ui = ui, server = server)
```

<script type="text/javascript" src="https://cdnjs.cloudflare.com/ajax/libs/iframe-resizer/3.5.16/iframeResizer.min.js"></script>
<style>
  iframe {
    min-width: 100%;
  }
</style>
<iframe id="myIframe" src="https://smartpop.shinyapps.io/ExploreAntenna/" scrolling="no" frameborder="no"></iframe>
<script>
  iFrameResize({
    heightCalculationMethod: 'taggedElement'
  });
</script>

### Type of BTS

After visual exploration several types of BTS can be identified:

* **Type 1**: BTS active all days with a regular pattern without noticeable anomalies.
* **Type 2**: BTS active all days with an irregular pattern and potential anomalies. 
* **Type 3**: BTS set up for special events, active for a couple of days (for the duration of the event) with a peak in SIM numbers (summer festivals...).
* **Type 4**: BTS covering commercial areas, most probably located inside a building with low range coverage as they show lower SIM numbers than BTS around.
* **Type 5**: Two or more BTS completing each other time series without overlap. They look like the same BTS but with different antennas
* **Type 5 bis**: same as type 5 but with overlap in time.
* **Type 6**: Two or more BTS having a regular patterns appearing simultaneously in the database. They look like newly installed BTS. With the new BTS, the number of SIM active in the BTS decrease since the users can connect to more BTS.
* **Type 7**: same as type 5 but without changed in activity of BTS around. It looks like an area that was previously poorly covered by the network and new users can access the network thanks to the new BTS. It concerns only few users otherwise some trends should be visible on the time series of total SIM over time.
* **Type 8**: BTS in events area/buildings (such as national stadium, spiroudome, Bruxelles expo) with high peak during irregular events.
* **Type 9**: BTS with SIM very close to 0 for each active days.

A lookup table is produced thanks to the shiny app to merge the location points (BTS / antenna) with activity complementing each other.
```{r, eval = F}
## Delete antenna with 0 activity

# Load lookup table
lookUp <- fread("/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Shiny/ExploreAntenna/antenna_discard_latlon.csv")
lookUp$NewID[lookUp$OldID == 938] <- 938 #Due to small error in the remaining of the analysis

# Remove antenna with low activity visually identified
Mobile_Phone <- Mobile_Phone[-which(Mobile_Phone$ID %in% na.omit(lookUp$`Antenne 0 (delete)`)),]
antenna_shp  <- antenna_shp[-which(antenna_shp$ID %in% na.omit(lookUp$`Antenne 0 (delete)`)), ] 

# Replace IDs in the full db and antenna shapefile
Mobile_Phone$ID[Mobile_Phone$ID %in% lookUp$OldID]  <- lookUp$NewID[na.omit(match(Mobile_Phone$ID, lookUp$OldID))]

lookUp2 <- lookUp[-which(lookUp$OldID == lookUp$NewID),]
antenna_shp  <- antenna_shp[-which(antenna_shp$ID %in% lookUp2$OldID), ] 

# Merge alls identical id (summarise)
library(dplyr)
Mobile_Phone <- Mobile_Phone %>%
                group_by(DATE_HOUR, ID) %>%
                dplyr::summarise(NBR_SIM_BE = sum(NBR_SIM_BE),
                          NBR_SIM_ROAMERS = sum(NBR_SIM_ROAMERS),
                          NBR_EVENTS_BE = sum(NBR_EVENTS_BE),
                          NBR_EVENTS_ROAMERS = sum(NBR_EVENTS_ROAMERS))
```

The final database is almost ready. Zero values are added to antenna with < 365 days of activity and specific areas (e.g. Bruxelles expo) identified during the visual exloration using the shiny app.

```{r, eval = F}
for (id in na.omit(lookUp$`event area (add 0)`)) {
  dates <- unique(Mobile_Phone$DATE_HOUR)[-which(unique(Mobile_Phone$DATE_HOUR) %in% Mobile_Phone$DATE_HOUR[Mobile_Phone$ID == id])]
  df <- data.frame(DATE_HOUR = dates,
                   ID = id,
                   NBR_SIM_BE = rep(0, length(dates)),
                   NBR_SIM_ROAMERS = rep(0, length(dates)),
                   NBR_EVENTS_BE = rep(0, length(dates)),
                   NBR_EVENTS_ROAMERS = rep(0, length(dates)))
  Mobile_Phone <- rbind.data.frame(Mobile_Phone, df)
}
```

{{% alert warning %}}
Assumption 3: There is *nobody* in areas covered by BTS located in "event areas"" (e.g. Bruxelles Expo) when these BTS are not active.
{{% /alert %}}

```{r, eval = F, echo = F}
fwrite(Mobile_Phone, file = "/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/CleanData/Mobile_Phone.csv")
writeOGR(antenna_shp, "/home/ubuntu/Dropbox/Research/Projects/SMARTPOP/Data/CleanData/antenna.shp", layer = "antenna", driver = "ESRI Shapefile")
```

Part 2 of the analysis is [here](https://damien-c-jacques.rbind.io/post/dynamic-population-mapping-smartpop-part-2/).