---
title: Nighttime Lights Calibration
author: Damien C. Jacques
summary: "This R Notebook describes the steps to calibrate DMSP-OLS Nighttime Lights Time Series using GDP data."
date: '2018-03-15'
slug: nighttime-lights-calibration
categories:
  - R
tags:
  - nighttime lights
  - remote sensing
---


## Plot Nighttime Lights Time Series for Senegal

First, we load the Version 4 DMSP-OLS Nighttime Lights Time Series (downloaded from [here](https://ngdc.noaa.gov/eog/dmsp/downloadV4composites.html)). 

```{r}
library(data.table)
library(raster)
library(rgdal)
library(usdm)

filenames <- list.files("/media/ubuntu/DATA/Data/Nightlight/", 
                        pattern = ".tif$", 
                        recursive = TRUE, 
                        full.names = T)[c(1,2,4:9,17:24,29:34)] # Only  select one raster for each year
filenames
r <- stack(filenames)
```

We plot the first layer to visualize the data.

```{r}
plot(r$F101992.v4b_web.stable_lights.avg_vis)
```

We load a shapefile of Senegal to set the zone of interest. 

```{r}
Senegal <- readOGR("/home/ubuntu/Dropbox/Research/Projects/Till_Poverty/Shapes_after_2008/LIMITE_CA_CR_2014_arr_ID.shp")
```

```{r}
plot(Senegal)
```

We will use the shapefile extent to crop the nighttime lights raster to the region of interest (to save computer memory). The first step is to set the `crs` of the different datasets to the same coordinate system. 

```{r}
crs(r)
crs(Senegal)
```

We use the crs of the raster stack (easier to change crs of small shapefile than large raster).

```{r}
Senegal <- spTransform(Senegal, crs(r))
```

Then, we crop the raster with the extent of the shapefile.

```{r}
r.zone <- crop(r, extent(Senegal))	

Senegal.r <- rasterize(Senegal, 
                       r.zone, 
                       field="ID", 
                       dataType = "INT1U") # Change dataType if nrow(shp) > 255 to INT2U or INT4U
r.zone[is.na(Senegal.r)] <- NA
```

```{r}
plot(r.zone[[1]])
plot(Senegal, add=T, lwd=0.5, lty=2)
```

We discard DN equal to 63 (due to saturation) and lower to 5 (to avoid blooming effect) (see for instance [Mukherjee et al. 2017](https://link.springer.com/article/10.1007/s40010-017-0444-8)).  

```{r}
r.zone[r.zone==63] <- NA
r.zone[r.zone < 5] <- NA
```

We plot the Sum Of Lights (SOL). We can see that the time series is not really smooth. This is partly due to a lack of onboard calibration.

```{r}
r.zone.sum <- cellStats(r.zone, sum)
plot(as.numeric(substr(names(r.zone), 4,7)), 
     r.zone.sum, 
     type="o", 
     xlab="Year", 
     ylab="Sum of Light (expressed in Digital Number)", tck = 0.02)
```


We will use GDP data of Senegal as validation/comparison data ([World Bank](https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?locations=SN)):

```{r}
GDP <- fread("/home/ubuntu/Dropbox/Research/Projects/Till_Poverty/GDP/API_NY.GDP.MKTP.CD_DS2_en_csv_v2/API_NY.GDP.MKTP.CD_DS2_en_csv_v2.csv")
GDP.Senegal <- GDP[GDP$`Country Name`=="Senegal"]
GDP.Senegal <- data.frame(year=as.numeric(colnames(GDP.Senegal[,5:62])),
                          GDP=as.numeric(GDP.Senegal[1,5:62]))
plot(GDP.Senegal$year, 
     GDP.Senegal$GDP, 
     type="o", xlab="year", ylab="GDP", tck = 0.02)
```

We can plot both time series on the same figure (scaling each variable) to compare the data.

```{r}
plot(as.numeric(substr(names(r.zone), 4,7)), 
     scale(r.zone.sum), 
     type="o", xlab="Year", 
     ylab="Sum of Light (expressed in Digital Number)", tck = 0.02)
lines(GDP.Senegal$year[GDP.Senegal$year%in%1992:2013], 
      scale(GDP.Senegal$GDP[GDP.Senegal$year%in%1992:2013]), 
      type="o", col="red")
legend("topleft", col=c("black", "red"), 
       lty=1, c("SOL", "GDP"), bty="n")
```
Correlation is already high but, as we will se, it will be improved after calibration. 

```{r}
cor(r.zone.sum, GDP.Senegal$GDP[GDP.Senegal$year%in%1992:2013])
```

## Inter-calibration between years

Pseudo-invariant features (PIFs) i.e., stable night-light features over time, are often used as references for calibrating the temporal images. Puerto Rico, Mauritius and Okinawa are used for calibration purpose. These regions have been selected by [Wu et al. 2013](https://www.tandfonline.com/doi/abs/10.1080/01431161.2013.820365) as invariant regions "based on the following considerations: first, these three regions are located in different geographic areas around the world and have worldwide representation; second, each region has a wide spread of digital number values from very low to very high, which would improve the accuracy of the intercalibration model; third, because all are far away from the mainland and are relatively isolated, they are rarely affected by night-time
light from other regions."


Let us first collect the data and crop the nighttime lights raster for each region of interest.

```{r}
PuertoRico <- readOGR("/home/ubuntu/Dropbox/Research/Projects/Till_Poverty/Nighlight/Shape_PFI/PFI/PuertoRico.shp")
Mauritius <- readOGR("/home/ubuntu/Dropbox/Research/Projects/Till_Poverty/Nighlight/Shape_PFI/PFI/MUS.shp")
Okinawa <- readOGR("/home/ubuntu/Dropbox/Research/Projects/Till_Poverty/Nighlight/Shape_PFI/PFI/Okinawa.shp")
par(mfrow=c(1,3))
plot(PuertoRico)
plot(Mauritius)
plot(Okinawa)

PuertoRico <- spTransform(PuertoRico, crs(r))
Mauritius <- spTransform(Mauritius, crs(r))
Okinawa <- spTransform(Okinawa, crs(r))

r.PuertoRico <- crop(r, extent(PuertoRico))	
r.Mauritius <- crop(r, extent(Mauritius))	
r.Okinawa <- crop(r, extent(Okinawa))	
```

We discard DN equal to 63 (due to saturation) and lower to 5 (to avoid blooming effect) (see for instance [Mukherjee et al. 2017](https://link.springer.com/article/10.1007/s40010-017-0444-8)).  

```{r}
r.PuertoRico[r.PuertoRico==63] <- NA
r.PuertoRico[r.PuertoRico < 5] <- NA

r.Mauritius[r.Mauritius==63] <- NA
r.Mauritius[r.Mauritius < 5] <- NA

r.Okinawa[r.Okinawa==63] <- NA
r.Okinawa[r.Okinawa < 5] <- NA


par(mfrow=c(1,3))
plot(r.PuertoRico[[1]])
plot(r.Mauritius[[1]])
plot(r.Okinawa[[1]])
```

Then we compute then standard deviation to select only the pixel with low temporal varability (NOT USED).

```{r}
# r.Mauritius.sd <- calc(r.Mauritius, sd)
# r.PuertoRico.sd <- calc(r.PuertoRico, sd)
# r.Okinawa.sd <- calc(r.Okinawa, sd)
# 
# par(mfrow=c(1,3))
# plot(r.Mauritius.sd)
# plot(r.PuertoRico.sd)
# plot(r.Okinawa.sd)
```

We use an arbitray threshold of a standard deviation equal to 10 (NOT USED).

```{r}
# r.PuertoRico.sub <- r.PuertoRico[r.PuertoRico.sd[,]<=10]
# r.Mauritius.sub <- r.Mauritius[r.Mauritius.sd[,]<=10]
# r.Okinawa.sub <- r.Okinawa[r.Okinawa.sd[,]<=10]
```

We then aggreggate the selected pixel time series into a matrix for further computation.

```{r}
r.select <- rbind(r.PuertoRico[,], r.Mauritius[,])
r.select <- rbind(r.select, r.Okinawa[,])
# r.select <- rbind(r.PuertoRico.sub, r.Mauritius.sub)
# r.select <- rbind(r.select, r.Okinawa.sub)
r.select <- r.select[complete.cases(r.select), ]
nrow(r.select)
```


Optimization procedure to find the best year of reference. We use a polynomial regression of degree 2 (quadratic model) to correct the time series.

```{r}
output <- data.frame(year=1992:2013, cor=1:22)
# 
# r3[r3==63] <- NA
# r3[r3 < 5] <- NA

for (i in 1:ncol(r.select)){
  print(i)
  r3 <- r.zone
  for (ii in 1:ncol(r.select)){
    # print(i)
    df <- data.frame(x=r.select[,ii], y=r.select[,i])
    model <- lm(y ~ x + I(x^2), data=df)
    # print(summary(model)$r.squared)
    # plot(df$x, df$y)
    # lines(df$x[order(df$x)],predict(model, df)[order(df$x)])
    name_layer <- names(r3[[ii]]) 
    names(r3[[ii]]) <- "x"
    r3[[ii]] <- raster::predict(r3[[ii]], model)
    names(r3[[ii]]) <- name_layer
  }
  r3[r3>63] <- 63
  r3[r3<0] <- 0
  r3.sum <- cellStats(r3, sum)
  output$cor[i] <- cor(r3.sum, 
                       GDP.Senegal$GDP[GDP.Senegal$year%in%as.numeric(substr(names(r.zone), 4,7))])
}
```

The year that maximize the correlation between GDP and SOL is used as reference:

```{r}
plot(output$year, output$cor, 
     ylab="Pearson Correlation", xlab="Year", tck=0.02, type="o")
```
```{r}
r.calibrated <- r.zone
for (ii in 1:ncol(r.select)){
  print(ii)
  df <- data.frame(x=r.select[,ii], y=r.select[,which.max(output$cor)])
  model <- lm(y ~ x + I(x^2), data=df)
  print(summary(model)$r.squared)
  # plot(df$x, df$y)
  # lines(df$x[order(df$x)],predict(model, df)[order(df$x)])
  name_layer <- names(r3[[ii]]) 
  names(r.calibrated[[ii]]) <- "x"
  r.calibrated[[ii]] <- raster::predict(r.calibrated[[ii]], model)
  names(r.calibrated[[ii]]) <- name_layer
}
r.calibrated[r.calibrated>63] <- 63
```

```{r}
r.calibrated.sum <- cellStats(r.calibrated, sum)
plot(as.numeric(substr(names(r.zone), 4,7)), 
     r.zone.sum, 
     type="o", xlab="Year", 
     ylab="Sum of Light (expressed in Digital Number)", 
     tck = 0.02, ylim=c(20000,90000))
lines(as.numeric(substr(names(r.zone), 4,7)), 
      r.calibrated.sum, 
      type="o", xlab="Year", 
      ylab="Sum of Light (expressed in Digital Number)",  
      ylim=c(20000,90000), lty=2)
legend("topleft", lty=1:2, c("Uncalibrated", "Calibrated"), bty="n")
```
```{r}
plot(as.numeric(substr(names(r.zone), 4,7)), 
     scale(r.calibrated.sum), 
     type="o", xlab="Year", ylab="SOL & GDP", tck = 0.02, lty=2)
lines(GDP.Senegal$year[GDP.Senegal$year%in%1992:2013], 
      scale(GDP.Senegal$GDP[GDP.Senegal$year%in%1992:2013]), type="o", col="red")
legend("topleft", col=c("black", "red"), 
       lty=2:1, c("SOL calibrated", "GDP"), bty="n")
```

```{r}
cor(r.calibrated.sum, GDP.Senegal$GDP[GDP.Senegal$year%in%1992:2013])
```

Export new calibrated raster:

```{r, eval = FALSE}
writeRaster(r.calibrated, 
            "/home/ubuntu/Dropbox/Research/Projects/Till_Poverty/Nighlight/Calibrated/NighttimeLights_calibratedV2.tif")
```




